### 2023-07-18 Rscripts/AMPHIBIAN.DATA.Rmd
hi this is the first commit! were uploading all the old codebase

---
title: "AMPHIBIAN SURVEYS"
author: "RAFAEL BAEZ"
date: "2022-10-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#dependencies
library(googlesheets4)
library(tidyverse)
library(vegan)
library(viridisLite)

#load in data sheet

```

## Project report

```{r data set up}
#load in Ohio data
MAIN<-read_sheet("https://docs.google.com/spreadsheets/d/1VIVJd4BpC5KS2o7APdG4W5Sf2H4hGTDMdPZ6oAZEL7E/edit#gid=0") %>% 
  arrange(Site)  %>%
  mutate(Beavers = as.logical(Beavers),
         Fish = as.logical(Fish),
         AMD = as.logical(Acid_Mine_Drainage)) %>%
  select(!Acid_Mine_Drainage)

#check structure of data sheet
str(MAIN)

```

```{r}
#DEMO: How to vegan package
###Diversity calculations----
#arrange data by site
SpeciesData <- MAIN %>%
  select(Site, Number_GreenFrogs, Number_AmericanBullfrogs, Number_EasternNewts, Number_AmericanToads, Number_NorthernSlimySalamander, Number_RavineSalamander, Number_GrayTreeFrog)%>%
  arrange(Site)

#make dataframe of species abundance data
SD <- SpeciesData %>%
  select(Number_GreenFrogs, Number_AmericanBullfrogs, Number_EasternNewts, Number_AmericanToads, Number_NorthernSlimySalamander, Number_RavineSalamander, Number_GrayTreeFrog)

#make abscence prescence table, calculate number of taxa per site
AbsPres <- SpeciesData %>%
  transmute(GrFr = case_when(Number_GreenFrogs > 0 ~ 1, TRUE ~ 0), 
     AmBu = case_when(Number_AmericanBullfrogs > 0 ~ 1, TRUE ~ 0), 
     EaNe = case_when(Number_EasternNewts > 0 ~ 1, TRUE ~ 0),
     AmTo = case_when(Number_AmericanToads > 0 ~ 1, TRUE ~ 0),
     NSlSa = case_when( Number_NorthernSlimySalamander > 0 ~ 1, TRUE ~ 0),
     RaSa = case_when(Number_RavineSalamander > 0 ~ 1, TRUE ~ 0),
     GrTrFr = case_when(Number_GrayTreeFrog > 0 ~ 1, TRUE ~ 0))

#it has the diversity now but i have to link it back to main to run my tests
MAIN<-MAIN%>% mutate("ShannonDiversity" =diversity(SD),
                     "SimpsonDiversity" = diversity(SD, index = 'simpson'),
                     "n_taxa" = AbsPres$n_taxa) %>%
  mutate("evenness" = ShannonDiversity/log(Richness))

MAIN
#now my main data frame has diversity at the end
#i must add values for richness 
#add values for average population size

###summary statistics----
#summary table of mean species counts and se species counts
Pop_summ <- SpeciesData %>%
  mutate(se_GF = sd(Number_GreenFrogs)/sqrt(8),
         se_AB = sd(Number_AmericanBullfrogs)/sqrt(8),
         se_EN = sd(Number_EasternNewts)/sqrt(8),
         se_AT = sd(Number_AmericanToads)/sqrt(8),
         se_NSS = sd(Number_NorthernSlimySalamander)/sqrt(8),
         se_RS = sd(Number_RavineSalamander)/sqrt(8),
         se_GTF = sd(Number_GrayTreeFrog)/sqrt(8)) %>%
  summarise(GF = mean(Number_GreenFrogs), se_GF = mean(se_GF), #calculate means
            AB =mean(Number_AmericanBullfrogs), se_AB = mean(se_AB),
            EN = mean(Number_EasternNewts), se_EN = mean(se_EN),
            AT = mean(Number_AmericanToads), se_AT = mean(se_AT),
            NSS = mean(Number_NorthernSlimySalamander), se_NSS = mean(se_NSS),
            RS = mean(Number_RavineSalamander), se_RS = mean(se_RS),
            GTF = mean(Number_GrayTreeFrog), se_GTF = mean(se_GTF))

str(Pop_summ)
#pivot summary stat table for graphing
#species order 
Pop_summ1 <- pivot_longer(Pop_summ, 
             cols = c(GF,AB,EN,AT,NSS, RS,GTF), names_to = "species")

```

```{r}
#diversity plots----
#shannon diversity index

#boxplots----

#linear model
Shannon_anova <- aov(data = MAIN, ShannonDiversity ~ Fish)
anova(Shannon_anova)

Simpson_aov <- aov(data = MAIN, SimpsonDiversity ~ Fish)
anova(Simpson_aov)
```

```{r}
#metamorphosis data
#green frogs as example
#calculate  post metamorphosis prop
MAIN <- MAIN %>%
  mutate(GF_pm_prop = Number_GreenFrogs_PostMetamorphosis/Number_GreenFrogs,
          AB_pm_prop = Number_AmericanBullfrogs_Postmetamorphosis/Number_AmericanBullfrogs,
         EN_pm_prop = Number_EasternNewts_PostMetamorphosis/Number_EasternNewts)

#plot of props
#green frogs
ggplot(MAIN, aes(x = AMD, y = GF_pm_prop, fill = AMD)) +
  geom_boxplot(outlier.size = 0) +
  geom_jitter(height = 0, width = 0.3, shape = 21, stroke = 0.7)

#American Bullfrogs
ggplot(MAIN, aes(x = AMD, y = AB_pm_prop, fill = AMD)) +
  geom_boxplot(outlier.size = 0) +
  geom_jitter(height = 0, width = 0.3, shape = 21, stroke = 0.7)

#eastern Newts
ggplot(MAIN, aes(x = AMD, y = EN_pm_prop, fill = AMD)) +
  geom_boxplot(outlier.size = 0) +
  geom_jitter(height = 0, width = 0.3, shape = 21, stroke = 0.7)

#stats tests
prop_test <- MAIN %>%
  select(AMD, Fish, Beavers, GF_pm_prop, AB_pm_prop, EN_pm_prop)

#green frog 
GF <- prop_test %>%
  select(AMD, GF_pm_prop) %>%
  mutate(rn = row_number()) %>%
  pivot_wider( names_from = AMD, values_from = GF_pm_prop)
AMD_T_GF <- na.omit(c(GF$`TRUE`))
AMD_F_GF <- na.omit(c(GF$`FALSE`))
t.test(x = AMD_T_GF, y = AMD_F_GF)

#American Bullfrog
AB <- prop_test %>%
  select(AMD, AB_pm_prop) %>%
  mutate(rn = row_number()) %>%
  pivot_wider( names_from = AMD, values_from = AB_pm_prop)
AMD_T_AB <- na.omit(c(AB$`TRUE`))
AMD_F_AB <- na.omit(c(AB$`FALSE`))
t.test(x = AMD_T_AB, y = AMD_F_AB)

```

![soda.](Photos/unnamed.jpg)

### 2023-07-18 Rscripts/biodiversity.R

@@ -1,0 +1,61 @@
+#dependencies----
+require(googlesheets4)
+require(tidyverse)
+require(vegan)
+
+#README:
+#this script reads the biodiveristy data from the 'Leaf Pack Project' sheet file on google drive @ "https://docs.google.com/spreadsheets/d/1_8Fph2bfNAaMCpI9q8LHModg78G5v4BVoOnJUHY-knE/edit#gid=0"
+#sourced by the controller.R script. 
+
+#data reforamtting----
+
+#data reforamtting----
+#read in sheets file from google drive link
+bd <- read_sheet("https://docs.google.com/spreadsheets/d/1_8Fph2bfNAaMCpI9q8LHModg78G5v4BVoOnJUHY-knE/edit#gid=1699683095",
+                 sheet = 1) %>%
+  filter(material != "Cellulose (Treated)") %>% #exclude Cellulose treated data from dataset
+  #reformat site names,attachment style to consistent standard values
+  mutate(site = str_remove(site, "-"), 
+         site = str_to_lower(site),
+         attachment_style = str_to_lower(attachment_style),
+         #calculate diversity indices for columns which start with an upper-case letter
+         shannon = diversity(select(., str_which(names(.), "^[:upper:]"))), 
+         simpson =diversity(select(., str_which(names(.), "^[:upper:]")), index = "simpson"),
+         invsimp = diversity(select(., str_which(names(.), "^[:upper:]")), index = "invsimpson"),
+         bag_quality = eval(as.numeric(str_extract(bag_quality, "[:digit:]"))/5)) %>% #convert bag quality to numeric
+  mutate(even = shannon/log(richness), #calcluate shannon evenness
+         mesh_size = case_when(material == "Biopolymer" ~ 10,
+                               material == "Cellulose" ~ 3, #derive mesh size value from material
+                               material == "Plastic" ~ 12.7,
+                               material == "Jute" ~ 17.78,
+                               material == "Cotton" ~ 5.08)) %>%
+  arrange(date)   #arrange by date
+
+#functional group catergorization----
+#create vectors of functional group member taxa
+collector <- c("MidgeFlies", "BlackFlies", "Planaria", "AquaticWorms", "Amphipods", "NetSpinningCaddisflies", "ClamsMussels")
+shredder <- c("Sowbugs", "CraneFlies", "Stoneflies", "RiffleBeetles")
[…]

### 2023-07-18 Rscripts/controller.R

@@ -1,0 +1,121 @@
+#dependencies----
+require(car)
+require(tidyverse)
+
+#source data processor scripts----
+source("Rscripts/biodiversity.R")#script to process biodiversity data
+source("Rscripts/water_quality.R")#script to process water quality data
+source("Rscripts/plot_source.R") #source code for custom plotting functions
+
+#join together summ_water_quality and biodiversity data
+joined <- full_join(x = wq_list$summ_water_quality, y = bd_list$biodiversity, by = c("date", "site"), multiple = "all")
+
+#statistical analyzes:
+#diversity model object list 
+#FORMULA: INDEX ~ MATERIAL + DATE
+diversity_mod_objects <- list(#taxa wide anova model objects
+                              shannon = aov(shannon ~ material + as.factor(date) , data = joined),
+                              simpson = aov(simpson ~ material + as.factor(date) , data = joined),
+                              invsimp = aov(simpson ~ material + as.factor(date) , data = joined),
+                              evenness = aov(even ~ material + as.factor(date) , data = joined),
+                              richness = aov(richness ~ material + as.factor(date), data = joined),
+                              #functional group anova model objects
+                              coll_shannon = aov(collector_shannon ~ material + as.factor(date)   , data = joined),
+                              scr_shannon = aov(shredder_shannon ~ material + as.factor(date)   , data = joined),
+                              pre_shannon = aov(predator_shannon ~ material + as.factor(date)   , data = joined),
+                              shr_shannon = aov(shredder_shannon ~ material + as.factor(date)   , data = joined))
+                
+#Type two anova list using car::Anova function 
+diversity_tests <- diversity_mod_objects %>%
+  {list(shannon = Anova(.$shannon),
+        simpson = Anova(.$simpson),
+        invsimp = Anova(.$invsimp),
+        eveneness = Anova(.$evenness),
+        richness = Anova(.$richness),
[…]

### 2023-07-18 Rscripts/deprecated_bagproj.R

@@ -1,0 +1,193 @@
+#README #README #README #README #README #README #README #README #README:
+#Hello! this is the data analysis script for the bag project. YOU MUST SOURCE bag_source.R IN ORDER FOR
+#THIS SCRIPT TO WORK!!!
+#control panel----
+#set to true to remove intermediate objects
+REMOVE_INTERMED = TRUE
+
+#reading in data----
+
+#this code is deprecated owo why are you snooping here???
+
+#BIODIVERSITY
+data <- read_sheet("https://docs.google.com/spreadsheets/d/1_8Fph2bfNAaMCpI9q8LHModg78G5v4BVoOnJUHY-knE/edit#gid=1699683095",
+                 sheet = 1) %>%
+  #reformat site names to match wq table
+  mutate(site = str_remove(site, "-"), #remove dashes
+     site = str_to_lower(site)) %>%
+  filter(material != "Cellulose (Treated)") %>%
+  #calculate base diversity indices, selecting for strings which start with an underscore 
+  mutate(shannon = diversity(select(., str_which(names(.), "^[:upper:]"))), 
+         simpson =diversity(select(., str_which(names(.), "^[:upper:]")), index = "simpson"),
+     invsimp = diversity(select(., str_which(names(.), "^[:upper:]")), index = "invsimpson"),
+   #calc numeric bag quality from string
+       bag_quality = eval(as.numeric(str_extract(bag_quality, "[:digit:]"))/5)) %>%
+  #calc evenness from shannon
+  mutate(even = shannon/log(richness),
+         #TODO: FILL OUT CASE_WHEN WITH THE REST OF THE MESH SIZE VALUES (ENSURE CORRECT SPELLING)
+         mesh_size = case_when(material == "Biopolymer" ~ 10,
+                               material == "Cellulose (Untreated)" ~ 3,
+                               material == "Plastic" ~ 12.7,
+                               material == "Jute" ~ 17.78,
+                               material == "Cotton" ~ 5.08)) %>%  
+  #arrange by date
+arrange(date)
+
+#JOINING WQ AND BIODIVERSITY
+#join water quality to data by date and site
+joined <- right_join(x = wq, y = data, by = c("date", "site"), multiple = "all") %>%
+select(material,date,site, everything())
+
+#FUNCTIONAL GROUPS 
+#vectors of funtional group column names
[…]

### 2023-07-18 Rscripts/plot_source.R

@@ -1,0 +1,84 @@
+#README #README #README #README #README #README #README #README #README:
+#Hello!  this is the source code used to create necessary functions for bag data analysis. in order for
+#the bag_project.R script to work, you must source this file first and foremost
+#DEPENDENCIES----
+require(vegan)
+require(ggpubr)
+require(tidyverse)
+require(googlesheets4)
+require(viridisLite)
+require(rlang)
+require(car)
+
+#STATISTICS----
+
+
+
+#PLOTTING----
+#set plotting theme function 
+theme_rafa <- function(base_size){
+  theme_pubr(base_family="sans") %+replace%
+    theme(panel.background  = element_blank(),
+          plot.background = element_rect(fill="transparent", colour=NA), 
+          legend.background = element_rect(fill="transparent", colour=NA),
+          legend.key = element_rect(fill="transparent", colour=NA),
+          legend.text = element_text(size=base_size * 0.7),
+          legend.title = element_text(size = base_size * 0.8, 
+                                      face = "bold"),
+          axis.text = element_text(size = base_size*0.85),
+          axis.title = element_text(size = base_size * 1.2),
+          plot.title.position = "panel",
+          plot.caption.position = "panel",
+          plot.title = element_text(
+            size = base_size * 1.25,
+            face = "bold", 
+            hjust = 0.5,
+            vjust = 1),
+          plot.subtitle = element_text(
+            size = base_size * 0.9,
+            vjust = 1),
+          plot.caption =  element_text(
+            size = base_size*0.75,
+            hjust = 0.5)
+    )
+}
+
+#bag type barplot function:
+#PARAMS: 
+#index(string): a string specifying the desired index
+bt_barplot <- function(index){
+  #calculate mean and standard error of chosen index
+  new_tbl <- joined%>%
+  filter(is.na(material) == FALSE)%>% #remove na values from material before calculations
+    group_by(material) %>% #group by material
+    mutate(index_se = sd(get(index), na.rm = T)/sqrt(n())) %>% # calc SE
+    summarise(index = mean(get(index), na.rm = T), index_se = mean(index_se)) #summarise into mean
+    
+  ggplot(new_tbl, aes(x = reorder(material, -index, sum), y = index)) +
+    geom_col(aes(fill = reorder(material, -index, sum))) +
[…]

### 2023-07-18 Rscripts/water_quality.R

@@ -1,0 +1,68 @@
+#dependencies:----
+require(tidyverse)
+require(googlesheets4)
+
+#README:
+#This script reads the water quality data from the 'Leaf Pack Project' sheet file on google drive @ "https://docs.google.com/spreadsheets/d/1_8Fph2bfNAaMCpI9q8LHModg78G5v4BVoOnJUHY-knE/edit#gid=0"
+#sourced by the controller.R script. 
+
+#data reforamtting----
+#read in sheets file from google drive link
+wq <- read_sheet("https://docs.google.com/spreadsheets/d/1_8Fph2bfNAaMCpI9q8LHModg78G5v4BVoOnJUHY-knE/edit#gid=0",
+                 sheet = 2, na = "NA") %>%
+  select(!c(d_mean, d_se, m_mean, m_se, u_mean, u_se))%>% #select out calculated columns
+  pivot_wider( names_from ="param", values_from = c("d_rep1", "d_rep2", "d_rep3",
+                                                    "m_rep1", "m_rep2", "m_rep3",
+                                                    "u_rep1", "u_rep2", "u_rep3")) #pivot wider  by the rep columns
+#separate into 3 tables corresp to site, and remove site prefix for easy merging
+#DOWNSTREAM
+d_wq <- wq %>%
+  select(date, str_which(names(wq), "^d")) %>%
+  mutate(site = "downstream") %>%
+  rename_with(.fn = ~ str_sub(.x, start = 3, end = -1), .cols = !date & !site)
+#MIDSTREAM
+m_wq <- wq %>%
+  select(date, str_which(names(wq), "^m")) %>%
+  mutate(site = "midstream") %>%
+  rename_with(.fn = ~ str_sub(.x, start = 3, end = -1), .cols = !date & !site)
+#UPSTREAM
+u_wq <- wq %>%
+  select(date, str_which(names(wq), "^u")) %>%
+  mutate(site = "upstream") %>%
+  rename_with(.fn = ~ str_sub(.x, start = 3, end = -1), .cols = !date & !site)
+#join all three tables together row-wise, place date and site at beginning, arrange by date
+wq <- bind_rows(d_wq,m_wq,u_wq) %>%
+  select(date, site, everything()) %>%
+  arrange(date)
+
+#pivot into longer format to condense rep columns into 1
+water_quality <- wq %>%
[…]

### 2023-07-19 Rscripts/biodiversity.R
Deleted junk files from directory added local copy of data xlsx file updated data reading code to use read_excel instead of read_sheets added PCA analysis to controller.R script

@@ -1,5 +1,5 @@
 #dependencies----
-require(googlesheets4)
+require(readxl)
 require(tidyverse)
 require(vegan)
 
@@ -11,8 +11,7 @@ require(vegan)
 
 #data reforamtting----
 #read in sheets file from google drive link
-bd <- read_sheet("https://docs.google.com/spreadsheets/d/1_8Fph2bfNAaMCpI9q8LHModg78G5v4BVoOnJUHY-knE/edit#gid=1699683095",
-                 sheet = 1) %>%
+bd <- read_excel("Rdata/LPP.xlsx" ,sheet = 1) %>%
   filter(material != "Cellulose (Treated)") %>% #exclude Cellulose treated data from dataset
   #reformat site names,attachment style to consistent standard values
   mutate(site = str_remove(site, "-"), 

### 2023-07-19 Rscripts/controller.R

@@ -7,8 +7,14 @@ source("Rscripts/biodiversity.R")#script to process biodiversity data
 source("Rscripts/water_quality.R")#script to process water quality data
 source("Rscripts/plot_source.R") #source code for custom plotting functions
 
-#join together summ_water_quality and biodiversity data
-joined <- full_join(x = wq_list$summ_water_quality, y = bd_list$biodiversity, by = c("date", "site"), multiple = "all")
+#join together summ_water_quality and biodiversity data, retaining only the obs in biodiversity
+joined <- right_join(x = wq_list$summ_water_quality, y = bd_list$biodiversity, by = c("date", "site"), multiple = "all")
+
+#principal component analysis:
+taxaCount_PCA <- prcomp(as.matrix(select(bd_list$biodiversity, MidgeFlies:RightHandedSnails)))
+#merge PCAs with rest of data to create full dataset
+LPP_FullData <- tibble(joined, #main dataset
+                       select(as.tibble(taxaCount_PCA$x), PC1,PC2,PC3))
 
 #statistical analyzes:
 #diversity model object list 

### 2023-07-19 Rscripts/water_quality.R

@@ -8,8 +8,7 @@ require(googlesheets4)
 
 #data reforamtting----
 #read in sheets file from google drive link
-wq <- read_sheet("https://docs.google.com/spreadsheets/d/1_8Fph2bfNAaMCpI9q8LHModg78G5v4BVoOnJUHY-knE/edit#gid=0",
-                 sheet = 2, na = "NA") %>%
+wq <- read_excel("Rdata/LPP.xlsx",sheet = 2, na = "NA") %>%
   select(!c(d_mean, d_se, m_mean, m_se, u_mean, u_se))%>% #select out calculated columns
   pivot_wider( names_from ="param", values_from = c("d_rep1", "d_rep2", "d_rep3",
                                                     "m_rep1", "m_rep2", "m_rep3",
@@ -47,6 +46,7 @@ water_quality <- wq %>%
   pivot_longer(cols = ends_with("_Flow"), values_to = "Flow",names_to = "rn_Flow") %>%
   select(!starts_with("rn_"))
 
+#NOTE TO SELF: stop calling across w/o argumetns or you might break your code!!!
 #summary statsitics table grouped by site
 wq_site_summary <- water_quality %>%
   group_by(site) %>%

### 2023-07-24 Rscripts/controller.R
adding correlation plot for PCAs, adding plot between PC1 and PC2

@@ -1,6 +1,8 @@
 #dependencies----
 require(car)
 require(tidyverse)
+require(corrplot)
+require(viridisLite)
 
 #source data processor scripts----
 source("Rscripts/biodiversity.R")#script to process biodiversity data
@@ -125,3 +127,10 @@ wq_list$site_summ_wq %>%
 arrange(desc(site)) %>%
 write_csv( file = "Rdata/wq_summary_by site.csv")
 
+
+corrplot(taxaCount_PCA$rotation[,1:3])
+
+ggplot(LPP_FullData, aes(x = PC1, y = PC2, fill = material)) +
+  geom_point(shape = 21) +
+  facet_wrap(~site) +
+  scale_fill_viridis_d()

### 2023-07-24 Rdocs/ManuscriptLPP.Rmd
deleted defunct data and folder added a markdown file for the manuscript repurposed old controller script as stats script added some dependencies to plot_source

---
title: Biodegradable alternatives to plastic mesh bags conventionally used in freshwater
  macroinvertebrate analysis within the Bronx River, NYC
author: "R.E. Segui-Baez, G.A. Rueda Moreno, R. Ward, K. Schneider Paolantonio"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
#knitr options----
knitr::opts_chunk$set(echo = TRUE)
#dependencies----
require(GGally)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r pre-requisite scripts, eval =FALSE}
#source data processor scripts----
source("Rscripts/biodiversity.R")#script to process biodiversity data
source("Rscripts/water_quality.R")#script to process water quality data
source("Rscripts/statistics.R") #script to analyze previously loaded data
source("Rscripts/plot_source.R") #source code for custom plotting functions

```

## Including Plots

You can also embed plots, for example:

```{r plots, echo=FALSE}
ggpairs(select(bd_list$biodiversity, MidgeFlies:Planaria))
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2023-07-24 Rscripts/plot_source.R

@@ -9,6 +9,8 @@ require(googlesheets4)
 require(viridisLite)
 require(rlang)
 require(car)
+require(corrplot)
+require(factoextra)
 
 #STATISTICS----
 

### 2023-07-24 Rscripts/statistics.R

@@ -1,13 +1,7 @@
 #dependencies----
 require(car)
+require(vegan)
 require(tidyverse)
-require(corrplot)
-require(viridisLite)
-
-#source data processor scripts----
-source("Rscripts/biodiversity.R")#script to process biodiversity data
-source("Rscripts/water_quality.R")#script to process water quality data
-source("Rscripts/plot_source.R") #source code for custom plotting functions
 
 #join together summ_water_quality and biodiversity data, retaining only the obs in biodiversity
 joined <- right_join(x = wq_list$summ_water_quality, y = bd_list$biodiversity, by = c("date", "site"), multiple = "all")
@@ -132,5 +126,9 @@ corrplot(taxaCount_PCA$rotation[,1:3])
 
 ggplot(LPP_FullData, aes(x = PC1, y = PC2, fill = material)) +
   geom_point(shape = 21) +
-  facet_wrap(~site) +
+  facet_wrap(~date) +
   scale_fill_viridis_d()
+
+#NMDS
+set.seed(420)
+ NMDS_LPP <- vegan::metaMDS(select(bd_list$biodiversity, MidgeFlies:RightHandedSnails))

### 2023-07-25 Rdocs/ManuscriptLPP.Rmd
moved plot source code to the Manuscript RMD file moved plots ot Manuscript RMD file standardized which data table is used in plot source code cleaned up depedency packages for each script

@@
 output: word_document
+bibliography: references.bib
 ---
@@
 #dependencies----
+require(tidyverse)
 require(GGally)
+require(vegan)
+require(ggpubr)
+require(tidyverse)
+require(viridisLite)
+require(rlang)
+require(car)
+require(corrplot)
+require(factoextra)
+
+#plotting source code----
+#set plotting theme function 
+theme_rafa <- function(base_size){
+  theme_pubr(base_family="sans") %+replace%
+    theme(panel.background  = element_blank(),
+          plot.background = element_rect(fill="transparent", colour=NA), 
+          legend.background = element_rect(fill="transparent", colour=NA),
+          legend.key = element_rect(fill="transparent", colour=NA),
+          legend.text = element_text(size=base_size * 0.7),
+          legend.title = element_text(size = base_size * 0.8, 
+                                      face = "bold"),
+          axis.text = element_text(size = base_size*0.85),
+          axis.title = element_text(size = base_size * 1.2),
+          plot.title.position = "panel",
+          plot.caption.position = "panel",
+          plot.title = element_text(
+            size = base_size * 1.25,
+            face = "bold", 
+            hjust = 0.5,
+            vjust = 1),
+          plot.subtitle = element_text(
+            size = base_size * 0.9,
+            vjust = 1),
+          plot.caption =  element_text(
+            size = base_size*0.75,
+            hjust = 0.5)
+    )
+}
+
+#bag type barplot function:
+#PARAMS: 
+  #index(string): a string specifying the desired index
+
+bt_barplot <- function(index){
+  #calculate mean and standard error of chosen index
+  new_tbl <- LPP_FullData%>%
+  filter(is.na(material) == FALSE)%>% #remove na values from material before calculations
+    group_by(material) %>% #group by material
+    mutate(index_se = sd(get(index), na.rm = T)/sqrt(n())) %>% # calc SE
+    summarise(index = mean(get(index), na.rm = T), index_se = mean(index_se)) #summarise into mean
+
+  ggplot(new_tbl, aes(x = reorder(material, -index, sum), y = index)) +
+    geom_col(aes(fill = reorder(material, -index, sum))) +
+    geom_errorbar(aes(ymin = index - 2*index_se, ymax = index + 2*index_se)) +
+
+    theme_bw() +
+    scale_fill_viridis_d(begin = 0.1, end = 0.9)+
+    theme_rafa(base_size = 15) +
+    theme(axis.text.x = element_text(angle = -15, vjust = 0.5))+
+    labs(x = "material", title = "Diversity Index by Bag material", y = paste(index),
+         fill = "material")
+}
+
+#water quality time series plots by site
+#PARAMS:
+  #param(string): a string specifying the water quality parameter
+
+wq_ts_BySite <- function(param){
+  #create symbol of parameter name and se for tidy eval in ggplot2 call
+  param<- sym(param)
+  param_se <- sym(paste(param,"se", sep = "_"))
+  ggplot(LPP_FullData, aes(x = date, y = !!param, color = site)) +
+    #geom layers
+    geom_point() +
+    geom_errorbar(aes(ymin = !!param - 2*!!param_se, ymax = !!param + 2*!!param_se)) + 
+    #theme adjustments
+    theme_bw() +
+    theme_rafa(base_size = 15) +
+    scale_color_viridis_d(begin = 0.1, end = 0.9)
+}
 ```
 
-## R Markdown
+##Abstract
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+##Introduction
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
+##Methods
+
+###Study Site and Timescale
+
+###Leaf Packs: Preparation
+
+###Leaf Packs: Retrieval and Sorting
+
+###Water Quality
+
+###Data Analyses
 
@@
 
-## Including Plots
+##Results
 
-You can also embed plots, for example:
+```{r result plots, echo=FALSE}
+#biodiversity plots----
+#cummul richness bag type barplots
+bt_barplot( "shannon")+ labs(x = "Material", y= "Shannon Diversity")
+bt_barplot( "simpson")+ labs(x = "Material",y= "Simpson Diversity")
+bt_barplot( "invsimp")+ labs(x = "Material",y= "Inverse-Simpson Diversity")
+bt_barplot( "richness")+ labs(x = "Material", y="Richness")
+bt_barplot("even")+ labs(x = "Material",y= "Shannon Evenness")
 
-```{r plots, echo=FALSE}
+#water quality dot plots
+#water quality by date on the y and site as color
+wq_ts_BySite("P")
+wq_ts_BySite("pH")
+wq_ts_BySite("NH3")
+wq_ts_BySite("DO")
+wq_ts_BySite("Temp")
+wq_ts_BySite("Cond")
+wq_ts_BySite("Nitrogen")
+wq_ts_BySite("Flow") + labs(title = "Flow rate seasonality",
+                            x = "Date", y = "Flow (m/s)")
 ggpairs(select(bd_list$biodiversity, MidgeFlies:Planaria))
@@
 
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+##Discussion
+
+##Acknowledgements
+
+##Referrences

### 2023-07-25 Rscripts/biodiversity.R

@@ -7,9 +7,9 @@ require(vegan)
 #this script reads the biodiveristy data from the 'Leaf Pack Project' sheet file on google drive @ "https://docs.google.com/spreadsheets/d/1_8Fph2bfNAaMCpI9q8LHModg78G5v4BVoOnJUHY-knE/edit#gid=0"
 #sourced by the controller.R script. 
 
-#data reforamtting----
+#data reformatting----
 
-#data reforamtting----
+#data reformatting----
 #read in sheets file from google drive link
 bd <- read_excel("Rdata/LPP.xlsx" ,sheet = 1) %>%
   filter(material != "Cellulose (Treated)") %>% #exclude Cellulose treated data from dataset
@@ -22,7 +22,7 @@ bd <- read_excel("Rdata/LPP.xlsx" ,sheet = 1) %>%
          simpson =diversity(select(., str_which(names(.), "^[:upper:]")), index = "simpson"),
          invsimp = diversity(select(., str_which(names(.), "^[:upper:]")), index = "invsimpson"),
          bag_quality = eval(as.numeric(str_extract(bag_quality, "[:digit:]"))/5)) %>% #convert bag quality to numeric
-  mutate(even = shannon/log(richness), #calcluate shannon evenness
+  mutate(even = shannon/log(richness), #calculate shannon evenness
          mesh_size = case_when(material == "Biopolymer" ~ 10,
                                material == "Cellulose" ~ 3, #derive mesh size value from material
                                material == "Plastic" ~ 12.7,
@@ -45,7 +45,7 @@ bd_matrices <- list(
   predator_matrix = as.matrix(bd[predator])
 )
 
-#add functional group diversity as collumn
+#add functional group diversity as column
 biodiversity <- bd %>%
   mutate(collector_shannon = diversity(bd_matrices[["collector_matrix"]]),
          shredder_shannon = diversity(bd_matrices[["shredder_matrix"]]),

### 2023-07-25 Rscripts/plot_source.R

@@ -1,86 +1,0 @@
-#README #README #README #README #README #README #README #README #README:
-#Hello!  this is the source code used to create necessary functions for bag data analysis. in order for
-#the bag_project.R script to work, you must source this file first and foremost
-#DEPENDENCIES----
-require(vegan)
-require(ggpubr)
-require(tidyverse)
-require(googlesheets4)
-require(viridisLite)
-require(rlang)
-require(car)
-require(corrplot)
-require(factoextra)
-
-#STATISTICS----
-
-
-
-#PLOTTING----
-#set plotting theme function 
-theme_rafa <- function(base_size){
-  theme_pubr(base_family="sans") %+replace%
-    theme(panel.background  = element_blank(),
-          plot.background = element_rect(fill="transparent", colour=NA), 
-          legend.background = element_rect(fill="transparent", colour=NA),
-          legend.key = element_rect(fill="transparent", colour=NA),
-          legend.text = element_text(size=base_size * 0.7),
-          legend.title = element_text(size = base_size * 0.8, 
-                                      face = "bold"),
-          axis.text = element_text(size = base_size*0.85),
-          axis.title = element_text(size = base_size * 1.2),
-          plot.title.position = "panel",
-          plot.caption.position = "panel",
-          plot.title = element_text(
-            size = base_size * 1.25,
-            face = "bold", 
-            hjust = 0.5,
-            vjust = 1),
-          plot.subtitle = element_text(
-            size = base_size * 0.9,
-            vjust = 1),
-          plot.caption =  element_text(
-            size = base_size*0.75,
-            hjust = 0.5)
-    )
-}
-
-#bag type barplot function:
-#PARAMS: 
-#index(string): a string specifying the desired index
-bt_barplot <- function(index){
-  #calculate mean and standard error of chosen index
-  new_tbl <- joined%>%
-  filter(is.na(material) == FALSE)%>% #remove na values from material before calculations
-    group_by(material) %>% #group by material
-    mutate(index_se = sd(get(index), na.rm = T)/sqrt(n())) %>% # calc SE
-    summarise(index = mean(get(index), na.rm = T), index_se = mean(index_se)) #summarise into mean
-    
-  ggplot(new_tbl, aes(x = reorder(material, -index, sum), y = index)) +
[…]

### 2023-07-25 Rscripts/statistics.R

@@ -1,6 +1,7 @@
 #dependencies----
 require(car)
 require(vegan)
+require(corrplot)
 require(tidyverse)
 
 #join together summ_water_quality and biodiversity data, retaining only the obs in biodiversity
@@ -13,19 +14,19 @@ LPP_FullData <- tibble(joined, #main dataset
                        select(as.tibble(taxaCount_PCA$x), PC1,PC2,PC3))
 
 #statistical analyzes:
-#diversity model object list 
+#diversity model object list---- 
 #FORMULA: INDEX ~ MATERIAL + DATE
 diversity_mod_objects <- list(#taxa wide anova model objects
-                              shannon = aov(shannon ~ material + as.factor(date) , data = joined),
-                              simpson = aov(simpson ~ material + as.factor(date) , data = joined),
-                              invsimp = aov(simpson ~ material + as.factor(date) , data = joined),
-                              evenness = aov(even ~ material + as.factor(date) , data = joined),
-                              richness = aov(richness ~ material + as.factor(date), data = joined),
+                              shannon = aov(shannon ~ material + as.factor(date) , data = LPP_FullData),
+                              simpson = aov(simpson ~ material + as.factor(date) , data = LPP_FullData),
+                              invsimp = aov(simpson ~ material + as.factor(date) , data = LPP_FullData),
+                              evenness = aov(even ~ material + as.factor(date) , data = LPP_FullData),
+                              richness = aov(richness ~ material + as.factor(date), data = LPP_FullData),
                               #functional group anova model objects
-                              coll_shannon = aov(collector_shannon ~ material + as.factor(date)   , data = joined),
-                              scr_shannon = aov(shredder_shannon ~ material + as.factor(date)   , data = joined),
-                              pre_shannon = aov(predator_shannon ~ material + as.factor(date)   , data = joined),
-                              shr_shannon = aov(shredder_shannon ~ material + as.factor(date)   , data = joined))
+                              coll_shannon = aov(collector_shannon ~ material + as.factor(date)   , data = LPP_FullData),
[…]

### 2023-07-25 Rscripts/water_quality.R

@@ -1,6 +1,5 @@
 #dependencies:----
 require(tidyverse)
-require(googlesheets4)
 
 #README:
 #This script reads the water quality data from the 'Leaf Pack Project' sheet file on google drive @ "https://docs.google.com/spreadsheets/d/1_8Fph2bfNAaMCpI9q8LHModg78G5v4BVoOnJUHY-knE/edit#gid=0"
@@ -50,7 +49,7 @@ water_quality <- wq %>%
 #summary statsitics table grouped by site
 wq_site_summary <- water_quality %>%
   group_by(site) %>%
-  mutate(across(.fns = ~sd(., na.rm = T)/n(), .names = "{.col}_se")) %>%
+  mutate(across(everything(), .fns = ~sd(., na.rm = T)/n(), .names = "{.col}_se")) %>%
   summarise(across( .fns = ~mean(., na.rm = TRUE)))
 
 #summary statsitics table grouped by site and date

### 2023-08-01 Rdocs/ManuscriptLPP.Rmd
Added almost all content from the google doc (missing  some of results, all of discussion) renamed Rplot folder to Rfigs for reasons

@@
 #set plotting theme function 
-theme_rafa <- function(base_size){
+theme_rafa <- function(base_size = 12){
   theme_pubr(base_family="sans") %+replace%
@@
   param_se <- sym(paste(param,"se", sep = "_"))
-  ggplot(LPP_FullData, aes(x = date, y = !!param, color = site)) +
+  ggplot(wq_list$raw_water_quality, aes(x = date, y = !!param, color = site)) +
     #geom layers
     geom_point() +
-    geom_errorbar(aes(ymin = !!param - 2*!!param_se, ymax = !!param + 2*!!param_se)) + 
+     geom_smooth(method = 'lm')+
     #theme adjustments
@@
 
+1.
+The traditional method of sampling aquatic systems utilizes plastic mesh leaf packs.
+Studies of amphibian and macroinvertebrate diversity, genetics, leaf degradation rates and bacterial colonization often utilize plastic mesh leaf packs within investigations.
+Due to factors such as inclement weather, debris accumulation, human error and symptoms of “urban stream syndrome”, these versatile packs with widespread usage are lost and contribute to large-scale pollution within aquatic systems.
+The presence of plastics in the Bronx River represents a threat to the organisms within the river system, with recent research suggesting the presence of twice as many microplastic particles as plankton at the mouth of the Bronx River.
+2.
+In this study, we tested four biodegradable alternative packs at three locations along the Bronx River at a site visited by New York University’s (NYU) Ecological Field Methods (EFM) course.
+We assessed these materials based on: durability, portability, longevity, affordability and effectiveness at collecting invertebrate data compared to the plastic.
+3.
+This study sampled 9,280 individuals from 17 discrete taxa.
+Amphipoda (genus Gammarus), Isopoda (genus Caecidotea), Diptera (suborder Nematocera), Basommatophora, and Oligochaetes made up the 5 most abundant taxa making up 95.8% of the total organisms seen.
+There were no statistically significant differences in terms of Shannon Diversity, Simpson Diversity, or Shannon Evenness between bag types but Richness was significant.
+The Jute and Biopolymer mesh performed exceptionally well in terms of durability, portability, longevity, affordability and effectiveness and surpassed the plastic standard in terms of richness.
+4.
+Our results presented several viable alternatives to the plastic standard, indicating a need for the scientific community to re-approach its methods for the analysis of ecosystemic health using leaf mesh packs.
+Results show that several non-plastic leaf packs do not impact the community assemblages, nor biodiversity of organisms within the packs.
+5.
+As climate change and urbanization increase with the passage of time, the local environments will have to contend with increasingly severe degrees of pollution and degradation.
+With this in mind, we encourage the utilization of novel methods that reduce the impact of scientific research on the environment and present a framework for future sustainable research in this area of inquiry.
+
 ##Introduction
+
+Despite being of critical ecological importance, urban streams are threatened by compounding anthropogenic stressors (Johnson et al.
+2020).
+Habitat disturbances, as facilitated by human activity, such as the creation of extensive impervious land cover and engineered conveyance networks, amplify the hydrological load being placed on streams (Groffman et al.
+2003).
+This amplification results in the degradation of both the physical, chemical, and biological stream environment, culminating in a phenomenon termed the ‘Urban Stream Syndrome’ (Meyer et al.
+2005).
+Although physical and chemical changes are generally consistent between regions, biological responses to urbanization are much more dependent on regional factors (Booth et al.
+2015).
+As such, the study of freshwater taxa and the wider ecological community has been critical for our understanding of the consequences of this Urban Stream Syndrome.
+To produce data for this kind of inquiry, the scientific community has looked to indicator species as a means of drawing conclusions about ecosystemic health; taxa of interest have ranged from frogs (Quasipaa boulengeri Yang 2022) to wood storks (Mycteria americana Frederick et al.
+2009) and even riparian lichen (Nelson 2015).
+Investigating aquatic macroinvertebrates as bioindicator species has been especially widespread during the examination of the health of aquatic ecosystems, being used across a variety of different biomes (Ilmonen et al 2009, Niba et al 2017, Arimoro et al 2021, Hu et al 2022).
+Macroinvertebrate taxa vary in their sensitivity and tolerance to environmental stressors, making them especially useful for these kinds of analyses (Pelletier et al 2010).
+The primary method of utilizing macroinvertebrates in biodiversity analysis within freshwater systems utilizes leaf mesh packs as the principal means of data collection (Merritt 1992, Robinson 2002, Kreutzweiser et al 2008, Vorste et al 2019).
+The usage of leaf mesh packs to study aquatic systems goes as far back as 1974 (Reice 1974) and they have been regularly used to investigate both macroinvertebrates and vertebrates alike (Waldron et al 2003, Talley et al 2012).
+This method employs plastic mesh bags filled with known quantities of leaves that are left in a stream and are allowed to be colonized by aquatic organisms for a standardized time after which they are retrieved and sorted.
+By creating a controlled and replicated habitat, leaf packs are efficient, cost effective and easy to use, frequently being deployed as school-group activities, advertised through organizations like the Stroud Leaf Pack Network as an initiative of the Stroud Water Research Center (https://stroudcenter.org/).
+However, due to inclement weather, shifts in current, water depth, debris accumulation, and a myriad of other factors, leaf packs can be lost during scientific proceedings (for example, see Meyer 1980).
+Thus complications begin to arise when one considers the wider scale of human activity within freshwater urban systems, specifically our relationship with plastic pollution.
+Leaf pack loss contributes to the introduction of both macroplastics and microplastics into aquatic systems.
+Defined as plastic particles less than 5mm in size, microplastics are often overlooked and are extremely hazardous to a range of organisms, including humans.
+Exposure to 3D printed plastics has been shown to induce a reduction in spontaneous embryonic movement, retardation of locomotion, and progressive paralysis in zebrafish (Walpitagama et al 2019).
+Once in water, plastics easily make their way into food webs, as observed in studies of plankton (Botterell et al 2022), lobsters (Welden and Cowie 2016), oysters (Sussarellu et al.
+2016), shrimp (Curren et al 2020), fish (Mercogliango et al 2020), birds (Provencher et al.
+2018, Carlin et al 2020) and orcas (Jepson et al 2016).
+In a recent literature review, Sequira et al.
+2020 reported that a median 60% of fish from 24 countries carry microplastics in their organs.
+The study reported even higher levels in carnivores, thus underlining the threat of bioaccumulation (Sequira et al.
+2020).
+If microplastic-contaminated fish are used for fish meal (Hanachi et al.
+2019) or consumed directly, they can pose a direct threat to human health (Naik 2019).
+The toxicity of microplastics is not limited to their presence in food webs: due to their generally positive charge and surface area, microplastics can also absorb contaminants like heavy metals (Chen et al 2022), PCBs, pesticides, and flame retardants (Othman et al 2021).
+They can also be colonized by bacteria - including some that can be pathogenic to fish (Viršek et al.
+2017).
+In order to lessen the impact of scientific studies on the ecosystems in question, novel methodologies must be developed which limit the use of plastics.
+The aim of this study is to determine alternatives to the plastic mesh bags conventionally used in studies of freshwater aquatic systems.
+We test four alternative materials and consider differences in their efficiency, cost, ease of implementation, and durability.
+Our aim is to promote the usage of these biodegradable alternatives such that we, as a community with the intention of examining the structure of a vulnerable system, can be conscientious of our contribution to the system’s degradation and choose more sustainable methods.
 
@@
 
+ The Bronx River is 23 miles long, stretching from Westchester County to the East River.
+The Bronx River Watershed is home to 2,451,404 people, many of whom come from minority groups and other underrepresented, displaced, and disenfranchised populations (Finewood et al 2023).
+ Our site in Shoelace Park is located approximately 9 km from the mouth of the river in Soundview Park in the South Bronx.
+It sits East of Woodlawn Cemetery, bordered by the Bronx River Parkway, a major highway on Shoelace Park’s west side, followed by a set of train tracks, a park buffer to the east, and surrounded by residential and commercial development.
+Shoelace Park exemplifies many urban stream systems (Fig 1).
+Additionally, this stream has been visited annually for
+several years during New York University’s Ecological Field Methods course.
+
+![Fig 1: The location of our study sites in Shoelace Park, Bronx NY.](Rfigs/INSERT_FIG1_HERE)
+
+We sampled three locations (termed upstream, midstream and downstream) for four rounds of sampling between 2022 and 2023 (Oct 8 - Nov 4 2022 termed “Nov retrieval”, Nov 4 - Dec 4, termed “Dec retrieval”, March 4 - March 26, termed “March retrieval” and March 26 - April 16, termed “April retrieval”).
+The three locations used during the study were between 30 and 40 meters apart (the locations were subject to minor changes in accessibility due to weather and changes in attachment points that varied with seasonal shifts of the landscape).
+
 ###Leaf Packs: Preparation
+
+During the course of the experiment, we followed procedures for placing, retrieving and analyzing leaf packs that have been used for decades within the Bronx River.
+These procedures were originally followed by Bronx River Restoration, a group that began cleaning up the Bronx River in the 1970s, and which subsequently transformed into the Bronx River Alliance, a non-profit and a division of the Bronx Parks.
+Because the Bronx River is geographically close to the Hudson River watershed, procedures recommended for use in the Hudson watershed were adopted for use in the Bronx River.
+Materials were selected according to their potential durability and lifespan, cost, and ease of use (our rationale can be seen in Appendix 1).
+In addition to the conventional plastic, four other materials were chosen to represent a wide variety of potential biodegradable alternatives: biopolymer, cellulose, jute and cotton (Table 1).
+Consistent with traditional protocol (Stroud Water Research Center), each bag was each filled with 30 grams of leaves, evenly divided between oak, maple and beech, that were collected near the sampling site and dried prior to use.
+Bags were closed by either knotting the material or weaving with nylon cord (see Appendix 2).
+During each sampling interval, each bag was anchored to exposed roots, submerged organic debris, trees or human constructed objects at the three sites in the river (Upstream, Midstream, and Downstream) using bricks and nylon cord.
+To account for seasonal changes in river depth, a generous amount of nylon was used to secure the bag to a brick which was then attached to submerged anthropogenic debris such as old pipes and human constructed structures, large exposed roots, or submerged organic debris like the fallen trunks of trees.
+The distance between the bag and the brick was minimized to stabilize the bag and reduce stress on the anchor knot.
+To allow for colonization by bacteria and invertebrates, bags remained submerged for four weeks (Stroud Water Research Center).
+
+```{r materials table}
+
+```
 
@@
 
+After four weeks, bags were removed by carefully cutting the bag from the brick and placing it into a submerged Ziplock bag.
+The ziplock bags were held in buckets and transported to the sampling station.
+Leaf packs were sorted in a manner consistent with the Stroud Protocol (Stroud Water Research Center).
+Bags were scored for durability on a 1 to 5 scale, with 5 describing bags being in roughly the same condition as when they were placed in the water and 1 being complete or near complete disintegration.
+Invertebrate identifications were made with laminated taxonomic keys and confirmed by photography or expert knowledge.
+Each individual observed was placed into its own temporary container alongside other members of its taxa, tallied up and recorded in data sheets.
+Taxa were classified by biotic health status and feeding status.
+After sampling, each organism was carefully transported back to the site along the river where it was retrieved.
+
 ###Water Quality
 
+Water quality parameters were recorded at each site during both the submerging and retrieval process.
+Nitrate (as Nitrogen) levels were measured using a Hach Nitraver 5 Nitrate Reagent test.
+Phosphate levels were measured using a Hach PhosVer 3 Phosphate Reagent test.
+Readings of Ammonia and pH were obtained using an API Aquarium Pharmaceuticals Ammonia (NH3/NH4+) and API high range pH test kits.
+Conductivity, dissolved oxygen and temperature were measured using YSI Ecosense EC 300 conductivity meter and probe and YSI EcoSense DO200-4/10 Dissolved Oxygen Probe.
+Flow rate (m/s) was determined using a ping pong ball and a meter stick.
+For most tests, we performed three trials at each site.
+
 ###Data Analyses
+All data analyses were performed using R, version 4.1.2 (R core team, 2021), and the vegan package was used to compute diversity indices (Oksanen et al. 2022)
 
@@
 
-```{r result plots, echo=FALSE}
-#biodiversity plots----
-#cummul richness bag type barplots
-bt_barplot( "shannon")+ labs(x = "Material", y= "Shannon Diversity")
-bt_barplot( "simpson")+ labs(x = "Material",y= "Simpson Diversity")
-bt_barplot( "invsimp")+ labs(x = "Material",y= "Inverse-Simpson Diversity")
-bt_barplot( "richness")+ labs(x = "Material", y="Richness")
-bt_barplot("even")+ labs(x = "Material",y= "Shannon Evenness")
+###Water Quality
 
+```{r wq plots}
 #water quality dot plots
@@
                             x = "Date", y = "Flow (m/s)")
+```
+
+ANOVA testing of water quality parameters showed statistically significant differences both across sampling date ( P < 0.001)  and site (P < 0.001).
+
+```{r result plots, echo=FALSE}
+#biodiversity plots----
+#cummul richness bag type barplots
+
 ggpairs(select(bd_list$biodiversity, MidgeFlies:Planaria))
+
+#PCA plot:----
+#dotplot between pc1 and pc2, with % explained variance annotated in axis labels
+ggplot(data = LPP_FullData, aes(x = PC1, y = PC2, fill = material)) +
+  geom_point(shape = 21, size = 2.5) + 
+geom_abline(slope = 0, intercept = 0, linetype = 2) +
+  geom_vline(xintercept = 0, linetype = 2) +
+
+  #thematics
+  scale_fill_viridis_d(begin = 0.1, end = 0.9, option = "G") +
+  theme_rafa(base_size = 16)
+
+#screeplot----
+fviz_eig(taxaCount_PCA) + theme_rafa(base_size = 16)
+```
+
+###Leaf Pack Efficiency
+####Overall Counts
+
+This study sampled 9,280 individuals from 17 discrete taxa.
+Amphipoda (genus Gammarus), Isopoda (genus Caecidotea), Diptera (suborder Nematocera), Basommatophora, and Oligochaetes made up the 5 most abundant taxa making up 95.8% of the total organisms seen.
+
+####Richness
+
+Taxonomic richness between bag types was statistically significant (P < 0.001, Figure 2).
+A subsequent Tukey-Kramer test revealed statistically significant differences between Cellulose and Biopolymer (P < 0.0001), Cellulose and Jute (P < 0.01), and Cellulose and Plastic (P < 0.0001).
+
+```{r richness plot}
+bt_barplot( "richness")+ labs(x = "Material", y="Richness")
+```
+
+####Diversity Indices and Evenness
+
+There were no statistically significant differences across bag types and sampling date using Simpson’s diversity index (P > 0.05)  and Shannon’s index (P > 0.05) (Figure 3).
+Both indices were examined to account for the influence of dominant species in the leaf packs.
+To determine if bag materials varied in their affinity for certain taxa (i.e.
+attracting/excluding organisms of specific functional groups), we also tested Shannon Evenness.
+ANOVA testing of Shannon Evenness by bag type resulted in marginally insignificant results (0.05 < P < 0.07) (Figure 4).
+These data can be found in Table 3.
+
+```{rshannon and simpson plots, fig.show="hold", out.width="50%"}
+bt_barplot( "shannon")+ labs(x = "Material", y= "Shannon Diversity")
+bt_barplot( "simpson")+ labs(x = "Material",y= "Simpson Diversity")
+```
+
+```{r inverse simpson and shannon evenness plots}
+bt_barplot( "invsimp")+ labs(x = "Material",y= "Inverse-Simpson Diversity")
+bt_barplot("even")+ labs(x = "Material",y= "Shannon Evenness")
 ```

### 2023-08-01 Rdocs/ManuscriptLPP.Rmd
adding materials and water quality tables

@@
 }
+```
+
+```{r pre-requisite scripts, echo=FALSE}
+#source data processor scripts----
+source("Rscripts/biodiversity.R")#script to process biodiversity data
+source("Rscripts/water_quality.R")#script to process water quality data
+source("Rscripts/statistics.R") #script to analyze previously loaded data
+source("Rscripts/plot_source.R") #source code for custom plotting functions
+
 ```
@@
 ```{r materials table}
-
+readxl::read_xlsx("Rfigs/summaryFigures.xlsx", sheet = 1) %>%
+knitr::kable(caption = "Table 1: Cost and Mesh Size By Material")
 ```
@@
 
-```{r pre-requisite scripts, eval =FALSE}
-#source data processor scripts----
-source("Rscripts/biodiversity.R")#script to process biodiversity data
-source("Rscripts/water_quality.R")#script to process water quality data
-source("Rscripts/statistics.R") #script to analyze previously loaded data
-source("Rscripts/plot_source.R") #source code for custom plotting functions
-
-```
-
 ##Results
@@
 
-```{r wq plots}
+```{r wq table}
+readxl::read_xlsx("Rfigs/summaryFigures.xlsx", sheet = 2) %>%
+knitr::kable(caption = "Table 2: Water quality parameters averaged over the course of the sampling period")
+```
+
+```{r wq plots, eval=FALSE}
 #water quality dot plots

### 2023-08-01 Rscripts/water_quality.R

@@ -45,18 +45,17 @@ water_quality <- wq %>%
   pivot_longer(cols = ends_with("_Flow"), values_to = "Flow",names_to = "rn_Flow") %>%
   select(!starts_with("rn_"))
 
-#NOTE TO SELF: stop calling across w/o argumetns or you might break your code!!!
-#summary statsitics table grouped by site
+#summary statistics table grouped by site
 wq_site_summary <- water_quality %>%
   group_by(site) %>%
   mutate(across(everything(), .fns = ~sd(., na.rm = T)/n(), .names = "{.col}_se")) %>%
   summarise(across( .fns = ~mean(., na.rm = TRUE)))
 
-#summary statsitics table grouped by site and date
+#summary statistics table grouped by site and date
 wq_full_summary <- water_quality %>%
   group_by(site, date) %>%
-  mutate(across(.fns = ~sd(., na.rm = T)/n(), .names = "{.col}_se")) %>%
-  summarise(across( .fns = ~mean(., na.rm = TRUE)))
+  mutate(across(everything(), .fns = ~sd(., na.rm = T)/n(), .names = "{.col}_se")) %>%
+  summarise(across(everything(), .fns = ~mean(., na.rm = TRUE)))
 
 #list of water quality objects
 wq_list <- list(raw_water_quality = water_quality,

### 2023-08-07 Rscripts/plot_source.R
Added Bray Curtis and the csv but im not sure how the working direcrory will work as this one is written to work through my desktop. However the csv is being pushed too so if it breaks im sure we can figure it out :/

@@ -82,3 +82,270 @@ wq_ts_BySite <- function(param){
     theme_rafa(base_size = 15) +
     scale_color_viridis_d(begin = 0.1, end = 0.9)
 }
+
+
+
+#BRAY CURTIS
+#trying to read in leaf pack paper from google sheet but downloaded as a csv to more quickly obtain a qorkable data table for r analysis
+
+
+
+getwd()
+setwd("/Users/reb/Desktop/ARE/LeafPackProject/Rdata")
+lpp<-read.csv("LPP.csv")
+#it worked but i have to put it in quotes
+
+#going to filter the lpp data frame so that it only has value from downstream and upstream
+
+#DU<-
+
+?mutate
+
+#creating data frame of lpp that will be mutated and chopped up
+Dstream<-lpp[lpp$site=="D", ]
+#this is all the data for sites with downstream trait
+
+Ustream<-lpp[lpp$site=="U",]
+#for upstream
+
+Mstream<-lpp[lpp$site=="M",]
+
+#so i have them seperated so now i can try to do the stat i want which is... i guess bray curtis but i need to have each table converted to be a sum of each organism so maybe i need to do column sums?
+
+
+#dropping the non numeric columns that I dont need
+
+MS<-subset(Mstream, select = -c(material,date ,bag_quality, site ,n_tol_taxa,n_ss_taxa, richness, n_s_taxa))
+
+US<-subset(Ustream, select = -c(material,date ,bag_quality, site ,n_tol_taxa,n_ss_taxa, richness, n_s_taxa))
+
+DS<-subset(Dstream, select = -c(material,date ,bag_quality, site ,n_tol_taxa,n_ss_taxa, richness, n_s_taxa))
+
+
+#creating new data frame with the species totals for each column 
+#proboem is i think i need to pivot the table so that the species names are columns instead of rows
+DSUMS<-as.data.frame(colSums(DS))
+
+#this switches the columns and rows
+DSUMS2<-as.data.frame(t(DSUMS))
+
+
+#now for midstream
+
+MSUMS<-as.data.frame(colSums(MS))
+
[…]

### 2023-08-09 Rscripts/statistics.R
cleaned up comments added scaling to PCA analysis

@@ -8,7 +8,8 @@ require(tidyverse)
 joined <- right_join(x = wq_list$summ_water_quality, y = bd_list$biodiversity, by = c("date", "site"), multiple = "all")
 
 #principal component analysis:
-taxaCount_PCA <- prcomp(as.matrix(select(bd_list$biodiversity, MidgeFlies:RightHandedSnails)))
+taxaCount_PCA <- prcomp(as.matrix(select(bd_list$biodiversity, MidgeFlies:RightHandedSnails)),
+                        scale. = T)
 #merge PCAs with rest of data to create full dataset
 LPP_FullData <- tibble(joined, #main dataset
                        select(as.tibble(taxaCount_PCA$x), PC1,PC2,PC3))
@@ -90,7 +91,7 @@ for(t in 1:length(wq_tests)){
 cor.test(LPP_FullData$bag_quality, LPP_FullData$richness)
 
 #muin::dredge for model selection
-#correl plot
+# bag quality  & richness correl plot
 ggplot(filter(LPP_FullData, is.na(material) == FALSE), aes(x = bag_quality*5, y = richness)) +
   geom_point(aes(color = material)) + geom_smooth(method = "lm", se = F, color = 'black')+
   theme_bw() + theme_rafa(base_size = 15) +
@@ -98,7 +99,7 @@ ggplot(filter(LPP_FullData, is.na(material) == FALSE), aes(x = bag_quality*5, y 
        title="Correlation Between Richness and Bag Quality Score ",
      caption =  "Pearson's r: 0.4212087 (P <0.01)")
 
-#output csv
+#output csv for writing kables
 wq_list$site_summ_wq %>%
 arrange(desc(site)) %>%
 write_csv( file = "Rdata/wq_summary_by site.csv")
@@ -106,11 +107,6 @@ write_csv( file = "Rdata/wq_summary_by site.csv")
 
 corrplot(taxaCount_PCA$rotation[,1:3])
 
-ggplot(LPP_FullData, aes(x = PC1, y = PC2, fill = material)) +
-  geom_point(shape = 21) +
-  facet_wrap(~date) +
-  scale_fill_viridis_d()
-
 #NMDS
-#set.seed(420)
-#NMDS_LPP <- vegan::metaMDS(select(bd_list$biodiversity, MidgeFlies:RightHandedSnails))
+set.seed(420)
+NMDS_LPP <- vegan::metaMDS(select(bd_list$biodiversity, MidgeFlies:RightHandedSnails))

### 2023-08-09 Rdocs/ManuscriptLPP.Rmd
cleaned up comments added scaling to PCA analysis
added PCA corrplot to ManuscriptLPP.Rmd

@@
 #PCA plot:----
+#prin comp correl plot for all components
+corrplot(taxaCount_PCA$rotation)
+
 #dotplot between pc1 and pc2, with % explained variance annotated in axis labels

### 2023-08-09 Rscripts/statistics.R

@@ -105,7 +105,7 @@ arrange(desc(site)) %>%
 write_csv( file = "Rdata/wq_summary_by site.csv")
 
 
-corrplot(taxaCount_PCA$rotation[,1:3])
+corrplot(taxaCount_PCA$rotation)
 
 #NMDS
 set.seed(420)

### 2023-08-11 Rdocs/ManuscriptLPP.Rmd
changes made during meeting with Manny

--- 2023-08-01 Rdocs/ManuscriptLPP.Rmd
@@
 ---
+
+```{r pre-requisite scripts, echo=FALSE}
+#source data processor scripts----
+source("Rscripts/biodiversity.R")#script to process biodiversity data
+source("Rscripts/water_quality.R")#script to process water quality data
+source("Rscripts/statistics.R") #script to analyze previously loaded data
+source("Rscripts/plot_source.R") #source code for custom plotting functions
+
+```
 
@@
   #index(string): a string specifying the desired index
-
 bt_barplot <- function(index){
+  #save index as character
+  char_index <- index
   #calculate mean and standard error of chosen index
@@
     geom_errorbar(aes(ymin = index - 2*index_se, ymax = index + 2*index_se)) +
+   # geom_text(aes(x = material, y = 0.5 + index, label = diversity_tukeys$char_index$groups$groups)) +
 
     theme_bw() +
-    scale_fill_viridis_d(begin = 0.1, end = 0.9)+
+    scale_fill_viridis_d()+
     theme_rafa(base_size = 15) +
@@
 }
-```
-
-```{r pre-requisite scripts, echo=FALSE}
-#source data processor scripts----
-source("Rscripts/biodiversity.R")#script to process biodiversity data
-source("Rscripts/water_quality.R")#script to process water quality data
-source("Rscripts/statistics.R") #script to analyze previously loaded data
-source("Rscripts/plot_source.R") #source code for custom plotting functions
-
 ```
@@
 #PCA plot:----
+#prin comp correl plot for all components
+corrplot(taxa_ct_PCA$rotation)
+
 #dotplot between pc1 and pc2, with % explained variance annotated in axis labels
@@
   geom_vline(xintercept = 0, linetype = 2) +
+  facet_wrap(~date) +
 
@@
 #screeplot----
-fviz_eig(taxaCount_PCA) + theme_rafa(base_size = 16)
+fviz_eig(taxa_ct_PCA) + theme_rafa(base_size = 16)
+
+ggplot(LPP_FullData, aes(x = sqrt(MidgeFlies))) +
+  geom_histogram()
 ```

### 2023-08-11 Rscripts/statistics.R

@@ -8,6 +8,10 @@ require(tidyverse)
 joined <- right_join(x = wq_list$summ_water_quality, y = bd_list$biodiversity, by = c("date", "site"), multiple = "all")
 
 #principal component analysis:
+taxa_ct_NMDS <- bd_list$biodiversity %>%
+  mutate(across(MidgeFlies:RightHandedSnails, ~log(.x + 1))) %>%
+  select(MidgeFlies:RightHandedSnails) %>%
+  prcomp(as.matrix(.), scale. = T)
 taxaCount_PCA <- prcomp(as.matrix(select(bd_list$biodiversity, MidgeFlies:RightHandedSnails)),
                         scale. = T)
 #merge PCAs with rest of data to create full dataset
@@ -44,13 +48,9 @@ diversity_tests <- diversity_mod_objects %>%
 #if the p values are below 95% confidence threshhold, perform tukey tests on them
 diversity_tukeys <- list()
 for(t in 1:length(diversity_tests)){
-  if(any(diversity_tests[[t]]$`Pr(>F)` < 0.05, na.rm = T) == TRUE)
-  {cat(names(diversity_tests[t]), "has P values below 0.05 threshhold, conducting tukey test", sep = " ")
-    diversity_tukeys[[t]] <- TukeyHSD(diversity_mod_objects[[t]])
+    diversity_tukeys[[t]] <- agricolae::HSD.test(diversity_mod_objects[[t]], "material", group = T)
     names(diversity_tukeys)[t] <- names(diversity_tests)[[t]]
-  }
-  #remove NA elements of tukey list
-  diversity_tukeys <- compact(diversity_tukeys)
+  
 }
 #WATER QUALITY----
 #wq model objects
@@ -93,8 +93,10 @@ cor.test(LPP_FullData$bag_quality, LPP_FullData$richness)
 #muin::dredge for model selection
 # bag quality  & richness correl plot
 ggplot(filter(LPP_FullData, is.na(material) == FALSE), aes(x = bag_quality*5, y = richness)) +
-  geom_point(aes(color = material)) + geom_smooth(method = "lm", se = F, color = 'black')+
-  theme_bw() + theme_rafa(base_size = 15) +
+  geom_jitter(aes(fill = material), shape =21) + 
+  scale_fill_viridis_d()+
+  theme_bw() +
+  theme_rafa(base_size = 16) +
   labs(x="Bag Quality Score", y="Richness",
[…]

### 2023-08-11 Rscripts/statistics.R
WIP

@@ -33,17 +33,17 @@ diversity_mod_objects <- list(#taxa wide anova model objects
                               pre_shannon = aov(predator_shannon ~ material + as.factor(date)   , data = LPP_FullData),
                               shr_shannon = aov(shredder_shannon ~ material + as.factor(date)   , data = LPP_FullData))
                 
-#Type two anova list using car::Anova function 
+#Type two anova list using car::car::Anova function 
 diversity_tests <- diversity_mod_objects %>%
-  {list(shannon = Anova(.$shannon),
-        simpson = Anova(.$simpson),
-        invsimp = Anova(.$invsimp),
-        eveneness = Anova(.$evenness),
-        richness = Anova(.$richness),
-        coll_shannon = Anova(.$coll_shannon),
-        scr_shannon = Anova(.$scr_shannon),
-        pre_shannon = Anova(.$pre_shannon),
-        shr_shannon = Anova(.$shr_shannon))}
+  {list(shannon = car::Anova(.$shannon),
+        simpson = car::Anova(.$simpson),
+        invsimp = car::Anova(.$invsimp),
+        eveneness = car::Anova(.$evenness),
+        richness = car::Anova(.$richness),
+        coll_shannon = car::Anova(.$coll_shannon),
+        scr_shannon = car::Anova(.$scr_shannon),
+        pre_shannon = car::Anova(.$pre_shannon),
+        shr_shannon = car::Anova(.$shr_shannon))}
 
 #if the p values are below 95% confidence threshhold, perform tukey tests on them
 diversity_tukeys <- list()

### 2023-08-11 Rscripts/stats.R

@@ -1,0 +1,38 @@
+#join diversity data and water quality data together
+joined <- right_join(x = wq_list$summ_water_quality, y = bd_list$biodiversity, by = c("date", "site"), multiple = "all")
+
+#statistical analyses----
+#DIVERSITY
+#FORMULA: INDEX ~ MATERIAL + DATE
+diversity_mod_objects <- list(#taxa wide anova model objects
+  shannon = aov(shannon ~ material + as.factor(date) , data = LPP_FullData),
+  simpson = aov(simpson ~ material + as.factor(date) , data = LPP_FullData),
+  invsimp = aov(simpson ~ material + as.factor(date) , data = LPP_FullData),
+  evenness = aov(even ~ material + as.factor(date) , data = LPP_FullData),
+  richness = aov(richness ~ material + as.factor(date), data = LPP_FullData),
+  #functional group anova model objects
+  coll_shannon = aov(collector_shannon ~ material + as.factor(date)   , data = LPP_FullData),
+  scr_shannon = aov(shredder_shannon ~ material + as.factor(date)   , data = LPP_FullData),
+  pre_shannon = aov(predator_shannon ~ material + as.factor(date)   , data = LPP_FullData),
+  shr_shannon = aov(shredder_shannon ~ material + as.factor(date)   , data = LPP_FullData))
+
+#Type two anova list using car::car::Anova function 
+diversity_tests <- diversity_mod_objects %>%
+  {list(shannon = car::Anova(.$shannon),
+        simpson = car::Anova(.$simpson),
+        invsimp = car::Anova(.$invsimp),
+        eveneness = car::Anova(.$evenness),
+        richness = car::Anova(.$richness),
+        coll_shannon = car::Anova(.$coll_shannon),
+        scr_shannon = car::Anova(.$scr_shannon),
+        pre_shannon = car::Anova(.$pre_shannon),
+        shr_shannon = car::Anova(.$shr_shannon))}
+
+#if the p values are below 95% confidence threshhold, perform tukey tests on them
+diversity_tukeys <- list()
+for(t in 1:length(diversity_tests)){
+  if(diversity_tests[[t]]){}
+  diversity_tukeys[[t]] <- agricolae::HSD.test(diversity_mod_objects[[t]], "material", group = T)
+  names(diversity_tukeys)[t] <- names(diversity_tests)[[t]]
+  
+}

### 2023-08-12 Rdocs/ManuscriptLPP.Rmd
Finalized changes to scripts for publication

@@
 source("Rscripts/water_quality.R")#script to process water quality data
-source("Rscripts/statistics.R") #script to analyze previously loaded data
+source("Rscripts/stats.R") #script to analyze previously loaded data
 source("Rscripts/plot_source.R") #source code for custom plotting functions

### 2023-08-12 Rscripts/biodiversity.R

@@ -1,26 +1,20 @@
 #dependencies----
-require(readxl)
 require(tidyverse)
-require(vegan)
-
-#README:
-#this script reads the biodiveristy data from the 'Leaf Pack Project' sheet file on google drive @ "https://docs.google.com/spreadsheets/d/1_8Fph2bfNAaMCpI9q8LHModg78G5v4BVoOnJUHY-knE/edit#gid=0"
-#sourced by the controller.R script. 
 
 #data reformatting----
 
 #data reformatting----
 #read in sheets file from google drive link
-bd <- read_excel("Rdata/LPP.xlsx" ,sheet = 1) %>%
+bd <- readxl::read_excel("Rdata/LPP.xlsx" ,sheet = 1) %>%
   filter(material != "Cellulose (Treated)") %>% #exclude Cellulose treated data from dataset
   #reformat site names,attachment style to consistent standard values
   mutate(site = str_remove(site, "-"), 
          site = str_to_lower(site),
          attachment_style = str_to_lower(attachment_style),
          #calculate diversity indices for columns which start with an upper-case letter
-         shannon = diversity(select(., str_which(names(.), "^[:upper:]"))), 
-         simpson =diversity(select(., str_which(names(.), "^[:upper:]")), index = "simpson"),
-         invsimp = diversity(select(., str_which(names(.), "^[:upper:]")), index = "invsimpson"),
+         shannon = vegan::diversity(select(., str_which(names(.), "^[:upper:]"))), 
+         simpson = vegan::diversity(select(., str_which(names(.), "^[:upper:]")), index = "simpson"),
+         invsimp = vegan::diversity(select(., str_which(names(.), "^[:upper:]")), index = "invsimpson"),
          bag_quality = eval(as.numeric(str_extract(bag_quality, "[:digit:]"))/5)) %>% #convert bag quality to numeric
   mutate(even = shannon/log(richness), #calculate shannon evenness
          mesh_size = case_when(material == "Biopolymer" ~ 10,
@@ -47,10 +41,10 @@ bd_matrices <- list(
 
 #add functional group diversity as column
 biodiversity <- bd %>%
-  mutate(collector_shannon = diversity(bd_matrices[["collector_matrix"]]),
-         shredder_shannon = diversity(bd_matrices[["shredder_matrix"]]),
-         scraper_shannon = diversity(bd_matrices[["scraper_matrix"]]),
[…]

### 2023-08-12 Rscripts/statistics.R

@@ -1,114 +1,0 @@
-#dependencies----
-require(car)
-require(vegan)
-require(corrplot)
-require(tidyverse)
-
-#join together summ_water_quality and biodiversity data, retaining only the obs in biodiversity
-joined <- right_join(x = wq_list$summ_water_quality, y = bd_list$biodiversity, by = c("date", "site"), multiple = "all")
-
-#principal component analysis:
-taxa_ct_NMDS <- bd_list$biodiversity %>%
-  mutate(across(MidgeFlies:RightHandedSnails, ~log(.x + 1))) %>%
-  select(MidgeFlies:RightHandedSnails) %>%
-  prcomp(as.matrix(.), scale. = T)
-taxaCount_PCA <- prcomp(as.matrix(select(bd_list$biodiversity, MidgeFlies:RightHandedSnails)),
-                        scale. = T)
-#merge PCAs with rest of data to create full dataset
-LPP_FullData <- tibble(joined, #main dataset
-                       select(as.tibble(taxaCount_PCA$x), PC1,PC2,PC3))
-
-#statistical analyzes:
-#diversity model object list---- 
-#FORMULA: INDEX ~ MATERIAL + DATE
-diversity_mod_objects <- list(#taxa wide anova model objects
-                              shannon = aov(shannon ~ material + as.factor(date) , data = LPP_FullData),
-                              simpson = aov(simpson ~ material + as.factor(date) , data = LPP_FullData),
-                              invsimp = aov(simpson ~ material + as.factor(date) , data = LPP_FullData),
-                              evenness = aov(even ~ material + as.factor(date) , data = LPP_FullData),
-                              richness = aov(richness ~ material + as.factor(date), data = LPP_FullData),
-                              #functional group anova model objects
-                              coll_shannon = aov(collector_shannon ~ material + as.factor(date)   , data = LPP_FullData),
-                              scr_shannon = aov(shredder_shannon ~ material + as.factor(date)   , data = LPP_FullData),
-                              pre_shannon = aov(predator_shannon ~ material + as.factor(date)   , data = LPP_FullData),
[…]

### 2023-08-12 Rscripts/stats.R

@@ -1,8 +1,7 @@
 #join diversity data and water quality data together
-joined <- right_join(x = wq_list$summ_water_quality, y = bd_list$biodiversity, by = c("date", "site"), multiple = "all")
+LPP_FullData <- right_join(x = wq_list$summ_water_quality, y = bd_list$biodiversity, by = c("date", "site"), multiple = "all")
 
-#statistical analyses----
-#DIVERSITY
+#DIVERSITY MODELS----
 #FORMULA: INDEX ~ MATERIAL + DATE
 diversity_mod_objects <- list(#taxa wide anova model objects
   shannon = aov(shannon ~ material + as.factor(date) , data = LPP_FullData),
@@ -28,11 +27,52 @@ diversity_tests <- diversity_mod_objects %>%
         pre_shannon = car::Anova(.$pre_shannon),
         shr_shannon = car::Anova(.$shr_shannon))}
 
-#if the p values are below 95% confidence threshhold, perform tukey tests on them
+#if the p values are below 95% confidence threshold, perform Tukey tests on them, else put NA
 diversity_tukeys <- list()
 for(t in 1:length(diversity_tests)){
-  if(diversity_tests[[t]]){}
-  diversity_tukeys[[t]] <- agricolae::HSD.test(diversity_mod_objects[[t]], "material", group = T)
-  names(diversity_tukeys)[t] <- names(diversity_tests)[[t]]
+  if(any(diversity_tests[[t]]$`Pr(>F)`[1:2] < 0.05)){
+    cat(paste(names(diversity_mod_objects[t]), "had significant (<0.05) results.",
+                "\n", "performing Tukey tests...",
+                "\n"))
+    diversity_tukeys[[t]] <- agricolae::HSD.test(diversity_mod_objects[[t]], "material", group = T)
+    names(diversity_tukeys)[t] <- names(diversity_tests)[[t]]
+  } else {
+    diversity_tukeys[[t]] <- NA
+  }
   
 }
+
+#WATER QUALITY MODELS----
+#FORMULA: PARAMETER ~ DATE + SITE
+#wq model objects
+wq_mod_objects <- list(
+  P = aov(P~as.factor(date)+site, data = wq_list$raw_water_quality),
+  pH = aov(pH~as.factor(date)+site, data = wq_list$raw_water_quality),
[…]

### 2023-08-12 Rscripts/water_quality.R

@@ -1,13 +1,6 @@
-#dependencies:----
-require(tidyverse)
-
-#README:
-#This script reads the water quality data from the 'Leaf Pack Project' sheet file on google drive @ "https://docs.google.com/spreadsheets/d/1_8Fph2bfNAaMCpI9q8LHModg78G5v4BVoOnJUHY-knE/edit#gid=0"
-#sourced by the controller.R script. 
-
 #data reforamtting----
 #read in sheets file from google drive link
-wq <- read_excel("Rdata/LPP.xlsx",sheet = 2, na = "NA") %>%
+wq <- readxl::read_excel("Rdata/LPP.xlsx",sheet = 2, na = "NA") %>%
   select(!c(d_mean, d_se, m_mean, m_se, u_mean, u_se))%>% #select out calculated columns
   pivot_wider( names_from ="param", values_from = c("d_rep1", "d_rep2", "d_rep3",
                                                     "m_rep1", "m_rep2", "m_rep3",

### 2023-08-21 Rscripts/deprecated_bagproj.R
deleting deprecated code

@@ -1,193 +1,0 @@
-#README #README #README #README #README #README #README #README #README:
-#Hello! this is the data analysis script for the bag project. YOU MUST SOURCE bag_source.R IN ORDER FOR
-#THIS SCRIPT TO WORK!!!
-#control panel----
-#set to true to remove intermediate objects
-REMOVE_INTERMED = TRUE
-
-#reading in data----
-
-#this code is deprecated owo why are you snooping here???
-
-#BIODIVERSITY
-data <- read_sheet("https://docs.google.com/spreadsheets/d/1_8Fph2bfNAaMCpI9q8LHModg78G5v4BVoOnJUHY-knE/edit#gid=1699683095",
-                 sheet = 1) %>%
-  #reformat site names to match wq table
-  mutate(site = str_remove(site, "-"), #remove dashes
-     site = str_to_lower(site)) %>%
-  filter(material != "Cellulose (Treated)") %>%
-  #calculate base diversity indices, selecting for strings which start with an underscore 
-  mutate(shannon = diversity(select(., str_which(names(.), "^[:upper:]"))), 
-         simpson =diversity(select(., str_which(names(.), "^[:upper:]")), index = "simpson"),
-     invsimp = diversity(select(., str_which(names(.), "^[:upper:]")), index = "invsimpson"),
-   #calc numeric bag quality from string
-       bag_quality = eval(as.numeric(str_extract(bag_quality, "[:digit:]"))/5)) %>%
-  #calc evenness from shannon
-  mutate(even = shannon/log(richness),
-         #TODO: FILL OUT CASE_WHEN WITH THE REST OF THE MESH SIZE VALUES (ENSURE CORRECT SPELLING)
-         mesh_size = case_when(material == "Biopolymer" ~ 10,
-                               material == "Cellulose (Untreated)" ~ 3,
-                               material == "Plastic" ~ 12.7,
-                               material == "Jute" ~ 17.78,
-                               material == "Cotton" ~ 5.08)) %>%  
-  #arrange by date
-arrange(date)
-
-#JOINING WQ AND BIODIVERSITY
-#join water quality to data by date and site
-joined <- right_join(x = wq, y = data, by = c("date", "site"), multiple = "all") %>%
-select(material,date,site, everything())
-
-#FUNCTIONAL GROUPS 
-#vectors of funtional group column names
[…]

### 2023-08-21 Rdocs/ManuscriptLPP.Rmd
fixed broken code in manuscript

@@
 source("Rscripts/stats.R") #script to analyze previously loaded data
-source("Rscripts/plot_source.R") #source code for custom plotting functions
 

### 2023-08-22 Rdocs/ManuscriptLPP.Rmd
added tukey groupings to diversity plots added bray curtis plots

@@
 bt_barplot <- function(index){
-  #save index as character
-  char_index <- index
+  #tukey tibble
+  groupings_tibble <- diversity_tukeys[[index]]$groups %>%
+rownames_to_column("material")
   #calculate mean and standard error of chosen index
@@
     summarise(index = mean(get(index), na.rm = T), index_se = mean(index_se)) #summarise into mean
+  #merge two tibbles
+  new_tbl <- inner_join(new_tbl, groupings_tibble, by ="material")
 
+  #plotting
   ggplot(new_tbl, aes(x = reorder(material, -index, sum), y = index)) +
@@
     geom_errorbar(aes(ymin = index - 2*index_se, ymax = index + 2*index_se)) +
-   # geom_text(aes(x = material, y = 0.5 + index, label = diversity_tukeys$char_index$groups$groups)) +
+geom_text(aes(y = (index + 2*index_se)+0.05, label =groups)) +
 
@@
 
+```{r}
+ggplot(BC_BySite_tibble, aes(x = site_comp, y = bc_score, fill = site_comp)) +geom_col() +
+  theme_rafa() + coord_cartesian(ylim = c(0, 1)) + scale_fill_viridis_d() +
+  labs(title = "Community Dissimilarity between sites",
+       x = "Site Comparisons", y =" Bray Curtis Dissimilarity", fill = "Comparisons")
+
+ggplot(BC_ByMat_tibble[7:10,], aes(x = mat_comp, y = bc_score, fill = mat_comp)) +geom_col() +
+  theme_rafa() + scale_fill_viridis_d()+
+  coord_cartesian(ylim = c(0, 1)) +
+  labs(title = "Community Dissimilairty between bag types", x = "Material Comparison",
+       y = "Bray Curtis Dissimilarity", fill = "Comparisons")
+```
+
 ##Discussion

### 2023-08-22 Rscripts/biodiversity.R

@@ -2,8 +2,6 @@
 require(tidyverse)
 
 #data reformatting----
-
-#data reformatting----
 #read in sheets file from google drive link
 bd <- readxl::read_excel("Rdata/LPP.xlsx" ,sheet = 1) %>%
   filter(material != "Cellulose (Treated)") %>% #exclude Cellulose treated data from dataset
@@ -45,6 +43,27 @@ biodiversity <- bd %>%
          shredder_shannon = vegan::diversity(bd_matrices[["shredder_matrix"]]),
          scraper_shannon = vegan::diversity(bd_matrices[["scraper_matrix"]]),
          predator_shannon = vegan::diversity(bd_matrices[["predator_matrix"]]))
+
+#Bray Curtis Dissimilarity calculation
+BD_BySite <- bd %>%
+  group_by(site) %>%
+  summarise(across(MidgeFlies:RightHandedSnails, sum)) 
+  BC_BySite <- vegdist(select(BD_BySite, !site)) 
+  BC_BySite_tibble <- tibble(
+    site_comp = c("down-midstream", "down-upstream", "mid-upstream"),
+    bc_score = as.vector(BC_BySite)
+  )
+  
+  
+  BD_ByMat <- bd %>%
+    group_by(material) %>%
+    summarise(across(MidgeFlies:RightHandedSnails, sum)) 
+  BC_ByMat <- vegdist(select(BD_ByMat, !material)) 
+  BC_ByMat_tibble <- tibble(
+    mat_comp = c("Biopolymer-cellulose", "BioPoly-cotton", "Cell-cotton", "Biopoly-Jute", "Cell-jute", "Cotton-Jute", "Biopolymer-Plastic", "Cell-Plastic", "Cotton-Plastic", "Jute-plastic"),
+    bc_score = as.vector(BC_ByMat)
+  )
+  
 
 #list of biodiversity objects
 bd_list <- list(biodiversity = biodiversity,

### 2023-08-22 Rscripts/stats.R

@@ -27,18 +27,15 @@ diversity_tests <- diversity_mod_objects %>%
         pre_shannon = car::Anova(.$pre_shannon),
         shr_shannon = car::Anova(.$shr_shannon))}
 
-#if the p values are below 95% confidence threshold, perform Tukey tests on them, else put NA
-diversity_tukeys <- list()
+#perform tukey tests for all indices
 for(t in 1:length(diversity_tests)){
-  if(any(diversity_tests[[t]]$`Pr(>F)`[1:2] < 0.05)){
+  
     cat(paste(names(diversity_mod_objects[t]), "had significant (<0.05) results.",
-                "\n", "performing Tukey tests...",
-                "\n"))
+              "\n", "performing Tukey tests...",
+              "\n"))
     diversity_tukeys[[t]] <- agricolae::HSD.test(diversity_mod_objects[[t]], "material", group = T)
     names(diversity_tukeys)[t] <- names(diversity_tests)[[t]]
-  } else {
-    diversity_tukeys[[t]] <- NA
-  }
+
   
 }
 
@@ -66,6 +63,7 @@ wq_tests <- wq_mod_objects %>%
   Cond = car::Anova(.$Cond),
   Flow = car::Anova(.$Flow))
   }
+
 
 #set seed for repeatability
 set.seed(420)

### 2023-09-13 Rdocs/ManuscriptLPP.Rmd
cleaning up directories, debugging biodiversity and stats scripts

@@
      geom_smooth(method = 'lm')+
+    #labels
+    labs(title = paste(param, "Seasonality"), x = "Date")+
     #theme adjustments
@@
 
-##Abstract
-
-1.
-The traditional method of sampling aquatic systems utilizes plastic mesh leaf packs.
-Studies of amphibian and macroinvertebrate diversity, genetics, leaf degradation rates and bacterial colonization often utilize plastic mesh leaf packs within investigations.
-Due to factors such as inclement weather, debris accumulation, human error and symptoms of “urban stream syndrome”, these versatile packs with widespread usage are lost and contribute to large-scale pollution within aquatic systems.
-The presence of plastics in the Bronx River represents a threat to the organisms within the river system, with recent research suggesting the presence of twice as many microplastic particles as plankton at the mouth of the Bronx River.
-2.
-In this study, we tested four biodegradable alternative packs at three locations along the Bronx River at a site visited by New York University’s (NYU) Ecological Field Methods (EFM) course.
-We assessed these materials based on: durability, portability, longevity, affordability and effectiveness at collecting invertebrate data compared to the plastic.
-3.
-This study sampled 9,280 individuals from 17 discrete taxa.
-Amphipoda (genus Gammarus), Isopoda (genus Caecidotea), Diptera (suborder Nematocera), Basommatophora, and Oligochaetes made up the 5 most abundant taxa making up 95.8% of the total organisms seen.
-There were no statistically significant differences in terms of Shannon Diversity, Simpson Diversity, or Shannon Evenness between bag types but Richness was significant.
-The Jute and Biopolymer mesh performed exceptionally well in terms of durability, portability, longevity, affordability and effectiveness and surpassed the plastic standard in terms of richness.
-4.
-Our results presented several viable alternatives to the plastic standard, indicating a need for the scientific community to re-approach its methods for the analysis of ecosystemic health using leaf mesh packs.
-Results show that several non-plastic leaf packs do not impact the community assemblages, nor biodiversity of organisms within the packs.
-5.
-As climate change and urbanization increase with the passage of time, the local environments will have to contend with increasingly severe degrees of pollution and degradation.
-With this in mind, we encourage the utilization of novel methods that reduce the impact of scientific research on the environment and present a framework for future sustainable research in this area of inquiry.
-
-##Introduction
-
-Despite being of critical ecological importance, urban streams are threatened by compounding anthropogenic stressors (Johnson et al.
-2020).
-Habitat disturbances, as facilitated by human activity, such as the creation of extensive impervious land cover and engineered conveyance networks, amplify the hydrological load being placed on streams (Groffman et al.
-2003).
-This amplification results in the degradation of both the physical, chemical, and biological stream environment, culminating in a phenomenon termed the ‘Urban Stream Syndrome’ (Meyer et al.
-2005).
-Although physical and chemical changes are generally consistent between regions, biological responses to urbanization are much more dependent on regional factors (Booth et al.
-2015).
-As such, the study of freshwater taxa and the wider ecological community has been critical for our understanding of the consequences of this Urban Stream Syndrome.
-To produce data for this kind of inquiry, the scientific community has looked to indicator species as a means of drawing conclusions about ecosystemic health; taxa of interest have ranged from frogs (Quasipaa boulengeri Yang 2022) to wood storks (Mycteria americana Frederick et al.
-2009) and even riparian lichen (Nelson 2015).
-Investigating aquatic macroinvertebrates as bioindicator species has been especially widespread during the examination of the health of aquatic ecosystems, being used across a variety of different biomes (Ilmonen et al 2009, Niba et al 2017, Arimoro et al 2021, Hu et al 2022).
-Macroinvertebrate taxa vary in their sensitivity and tolerance to environmental stressors, making them especially useful for these kinds of analyses (Pelletier et al 2010).
-The primary method of utilizing macroinvertebrates in biodiversity analysis within freshwater systems utilizes leaf mesh packs as the principal means of data collection (Merritt 1992, Robinson 2002, Kreutzweiser et al 2008, Vorste et al 2019).
-The usage of leaf mesh packs to study aquatic systems goes as far back as 1974 (Reice 1974) and they have been regularly used to investigate both macroinvertebrates and vertebrates alike (Waldron et al 2003, Talley et al 2012).
-This method employs plastic mesh bags filled with known quantities of leaves that are left in a stream and are allowed to be colonized by aquatic organisms for a standardized time after which they are retrieved and sorted.
-By creating a controlled and replicated habitat, leaf packs are efficient, cost effective and easy to use, frequently being deployed as school-group activities, advertised through organizations like the Stroud Leaf Pack Network as an initiative of the Stroud Water Research Center (https://stroudcenter.org/).
-However, due to inclement weather, shifts in current, water depth, debris accumulation, and a myriad of other factors, leaf packs can be lost during scientific proceedings (for example, see Meyer 1980).
-Thus complications begin to arise when one considers the wider scale of human activity within freshwater urban systems, specifically our relationship with plastic pollution.
-Leaf pack loss contributes to the introduction of both macroplastics and microplastics into aquatic systems.
-Defined as plastic particles less than 5mm in size, microplastics are often overlooked and are extremely hazardous to a range of organisms, including humans.
-Exposure to 3D printed plastics has been shown to induce a reduction in spontaneous embryonic movement, retardation of locomotion, and progressive paralysis in zebrafish (Walpitagama et al 2019).
-Once in water, plastics easily make their way into food webs, as observed in studies of plankton (Botterell et al 2022), lobsters (Welden and Cowie 2016), oysters (Sussarellu et al.
-2016), shrimp (Curren et al 2020), fish (Mercogliango et al 2020), birds (Provencher et al.
-2018, Carlin et al 2020) and orcas (Jepson et al 2016).
-In a recent literature review, Sequira et al.
-2020 reported that a median 60% of fish from 24 countries carry microplastics in their organs.
-The study reported even higher levels in carnivores, thus underlining the threat of bioaccumulation (Sequira et al.
-2020).
-If microplastic-contaminated fish are used for fish meal (Hanachi et al.
-2019) or consumed directly, they can pose a direct threat to human health (Naik 2019).
-The toxicity of microplastics is not limited to their presence in food webs: due to their generally positive charge and surface area, microplastics can also absorb contaminants like heavy metals (Chen et al 2022), PCBs, pesticides, and flame retardants (Othman et al 2021).
-They can also be colonized by bacteria - including some that can be pathogenic to fish (Viršek et al.
-2017).
-In order to lessen the impact of scientific studies on the ecosystems in question, novel methodologies must be developed which limit the use of plastics.
-The aim of this study is to determine alternatives to the plastic mesh bags conventionally used in studies of freshwater aquatic systems.
-We test four alternative materials and consider differences in their efficiency, cost, ease of implementation, and durability.
-Our aim is to promote the usage of these biodegradable alternatives such that we, as a community with the intention of examining the structure of a vulnerable system, can be conscientious of our contribution to the system’s degradation and choose more sustainable methods.
-
-##Methods
-
-###Study Site and Timescale
-
- The Bronx River is 23 miles long, stretching from Westchester County to the East River.
-The Bronx River Watershed is home to 2,451,404 people, many of whom come from minority groups and other underrepresented, displaced, and disenfranchised populations (Finewood et al 2023).
- Our site in Shoelace Park is located approximately 9 km from the mouth of the river in Soundview Park in the South Bronx.
-It sits East of Woodlawn Cemetery, bordered by the Bronx River Parkway, a major highway on Shoelace Park’s west side, followed by a set of train tracks, a park buffer to the east, and surrounded by residential and commercial development.
-Shoelace Park exemplifies many urban stream systems (Fig 1).
-Additionally, this stream has been visited annually for
-several years during New York University’s Ecological Field Methods course.
-
-![Fig 1: The location of our study sites in Shoelace Park, Bronx NY.](Rfigs/INSERT_FIG1_HERE)
-
-We sampled three locations (termed upstream, midstream and downstream) for four rounds of sampling between 2022 and 2023 (Oct 8 - Nov 4 2022 termed “Nov retrieval”, Nov 4 - Dec 4, termed “Dec retrieval”, March 4 - March 26, termed “March retrieval” and March 26 - April 16, termed “April retrieval”).
-The three locations used during the study were between 30 and 40 meters apart (the locations were subject to minor changes in accessibility due to weather and changes in attachment points that varied with seasonal shifts of the landscape).
-
-###Leaf Packs: Preparation
-
-During the course of the experiment, we followed procedures for placing, retrieving and analyzing leaf packs that have been used for decades within the Bronx River.
-These procedures were originally followed by Bronx River Restoration, a group that began cleaning up the Bronx River in the 1970s, and which subsequently transformed into the Bronx River Alliance, a non-profit and a division of the Bronx Parks.
-Because the Bronx River is geographically close to the Hudson River watershed, procedures recommended for use in the Hudson watershed were adopted for use in the Bronx River.
-Materials were selected according to their potential durability and lifespan, cost, and ease of use (our rationale can be seen in Appendix 1).
-In addition to the conventional plastic, four other materials were chosen to represent a wide variety of potential biodegradable alternatives: biopolymer, cellulose, jute and cotton (Table 1).
-Consistent with traditional protocol (Stroud Water Research Center), each bag was each filled with 30 grams of leaves, evenly divided between oak, maple and beech, that were collected near the sampling site and dried prior to use.
-Bags were closed by either knotting the material or weaving with nylon cord (see Appendix 2).
-During each sampling interval, each bag was anchored to exposed roots, submerged organic debris, trees or human constructed objects at the three sites in the river (Upstream, Midstream, and Downstream) using bricks and nylon cord.
-To account for seasonal changes in river depth, a generous amount of nylon was used to secure the bag to a brick which was then attached to submerged anthropogenic debris such as old pipes and human constructed structures, large exposed roots, or submerged organic debris like the fallen trunks of trees.
-The distance between the bag and the brick was minimized to stabilize the bag and reduce stress on the anchor knot.
-To allow for colonization by bacteria and invertebrates, bags remained submerged for four weeks (Stroud Water Research Center).
-
-```{r materials table}
-readxl::read_xlsx("Rfigs/summaryFigures.xlsx", sheet = 1) %>%
-knitr::kable(caption = "Table 1: Cost and Mesh Size By Material")
-```
-
-###Leaf Packs: Retrieval and Sorting
-
-After four weeks, bags were removed by carefully cutting the bag from the brick and placing it into a submerged Ziplock bag.
-The ziplock bags were held in buckets and transported to the sampling station.
-Leaf packs were sorted in a manner consistent with the Stroud Protocol (Stroud Water Research Center).
-Bags were scored for durability on a 1 to 5 scale, with 5 describing bags being in roughly the same condition as when they were placed in the water and 1 being complete or near complete disintegration.
-Invertebrate identifications were made with laminated taxonomic keys and confirmed by photography or expert knowledge.
-Each individual observed was placed into its own temporary container alongside other members of its taxa, tallied up and recorded in data sheets.
-Taxa were classified by biotic health status and feeding status.
-After sampling, each organism was carefully transported back to the site along the river where it was retrieved.
-
-###Water Quality
-
-Water quality parameters were recorded at each site during both the submerging and retrieval process.
-Nitrate (as Nitrogen) levels were measured using a Hach Nitraver 5 Nitrate Reagent test.
-Phosphate levels were measured using a Hach PhosVer 3 Phosphate Reagent test.
-Readings of Ammonia and pH were obtained using an API Aquarium Pharmaceuticals Ammonia (NH3/NH4+) and API high range pH test kits.
-Conductivity, dissolved oxygen and temperature were measured using YSI Ecosense EC 300 conductivity meter and probe and YSI EcoSense DO200-4/10 Dissolved Oxygen Probe.
-Flow rate (m/s) was determined using a ping pong ball and a meter stick.
-For most tests, we performed three trials at each site.
-
-###Data Analyses
-All data analyses were performed using R, version 4.1.2 (R core team, 2021), and the vegan package was used to compute diversity indices (Oksanen et al. 2022)
-
-##Results
-
-###Water Quality
+### Tables
 
@@
 
+### Plots
+
 ```{r wq plots, eval=FALSE}
-#water quality dot plots
+#water quality dot plots----
 #water quality by date on the y and site as color
-wq_ts_BySite("P")
+wq_ts_BySite("P") +labs( title = "Phosphorous Seasonality",y = "Phosphorous (mg/L)")
 wq_ts_BySite("pH")
-wq_ts_BySite("NH3")
-wq_ts_BySite("DO")
-wq_ts_BySite("Temp")
-wq_ts_BySite("Cond")
-wq_ts_BySite("Nitrogen")
-wq_ts_BySite("Flow") + labs(title = "Flow rate seasonality",
+wq_ts_BySite("NH3") +labs(y = "Ammonia (mg/L)")
+wq_ts_BySite("DO") + labs(title = "Dissolved Oxygen Seasonality",
+                          y = "% DO")
+wq_ts_BySite("Temp") + labs(title = "Temperature Seasonality",
+                            y = "Temp (°C)")
+wq_ts_BySite("Cond") + labs(title = "Conductivity Seasonality", y = "Conductivity (mS)")
+wq_ts_BySite("Nitrogen") + labs(y = "Nitrogen (mg/L)")
+wq_ts_BySite("Flow") + labs(title = "Flow rate Seasonality",
                             x = "Date", y = "Flow (m/s)")
-```
 
-ANOVA testing of water quality parameters showed statistically significant differences both across sampling date ( P < 0.001)  and site (P < 0.001).
-
-```{r result plots, echo=FALSE}
 #biodiversity plots----
-#cummul richness bag type barplots
-
-ggpairs(select(bd_list$biodiversity, MidgeFlies:Planaria))
-
-#PCA plot:----
-#prin comp correl plot for all components
-corrplot(taxa_ct_PCA$rotation)
-
-#dotplot between pc1 and pc2, with % explained variance annotated in axis labels
-ggplot(data = LPP_FullData, aes(x = PC1, y = PC2, fill = material)) +
-  geom_point(shape = 21, size = 2.5) + 
-geom_abline(slope = 0, intercept = 0, linetype = 2) +
-  geom_vline(xintercept = 0, linetype = 2) +
-  facet_wrap(~date) +
-
-#thematics
-  scale_fill_viridis_d(begin = 0.1, end = 0.9, option = "G") +
-  theme_rafa(base_size = 16)
-
-#screeplot----
-fviz_eig(taxa_ct_PCA) + theme_rafa(base_size = 16)
-
-ggplot(LPP_FullData, aes(x = sqrt(MidgeFlies))) +
-  geom_histogram()
-```
-
-###Leaf Pack Efficiency
-####Overall Counts
-
-This study sampled 9,280 individuals from 17 discrete taxa.
-Amphipoda (genus Gammarus), Isopoda (genus Caecidotea), Diptera (suborder Nematocera), Basommatophora, and Oligochaetes made up the 5 most abundant taxa making up 95.8% of the total organisms seen.
-
-####Richness
-
-Taxonomic richness between bag types was statistically significant (P < 0.001, Figure 2).
-A subsequent Tukey-Kramer test revealed statistically significant differences between Cellulose and Biopolymer (P < 0.0001), Cellulose and Jute (P < 0.01), and Cellulose and Plastic (P < 0.0001).
-
-```{r richness plot}
 bt_barplot( "richness")+ labs(x = "Material", y="Richness")  
-```
-
-####Diversity Indices and Evenness
-
-There were no statistically significant differences across bag types and sampling date using Simpson’s diversity index (P > 0.05)  and Shannon’s index (P > 0.05) (Figure 3).
-Both indices were examined to account for the influence of dominant species in the leaf packs.
-To determine if bag materials varied in their affinity for certain taxa (i.e.
-attracting/excluding organisms of specific functional groups), we also tested Shannon Evenness.
-ANOVA testing of Shannon Evenness by bag type resulted in marginally insignificant results (0.05 < P < 0.07) (Figure 4).
-These data can be found in Table 3.
-
-```{rshannon and simpson plots, fig.show="hold", out.width="50%"}
 bt_barplot( "shannon")+ labs(x = "Material", y= "Shannon Diversity")
 bt_barplot( "simpson")+ labs(x = "Material",y= "Simpson Diversity")
-```
-
-```{r inverse simpson and shannon evenness plots}
 bt_barplot( "invsimp")+ labs(x = "Material",y= "Inverse-Simpson Diversity")
 bt_barplot("even")+ labs(x = "Material",y= "Shannon Evenness")
-```
 
-```{r}
+#Bray Curtis Dissimil plots----
+#by site
 ggplot(BC_BySite_tibble, aes(x = site_comp, y = bc_score, fill = site_comp)) +geom_col() +
@@
 
+#by material (plastic comparisons only)
 ggplot(BC_ByMat_tibble[7:10,], aes(x = mat_comp, y = bc_score, fill = mat_comp)) +geom_col() +
@@
        y = "Bray Curtis Dissimilarity", fill = "Comparisons")
+
 ```
-
-##Discussion
-
-##Acknowledgements
-
-##Referrences

### 2023-09-13 Rscripts/biodiversity.R

@@ -44,21 +44,22 @@ biodiversity <- bd %>%
          scraper_shannon = vegan::diversity(bd_matrices[["scraper_matrix"]]),
          predator_shannon = vegan::diversity(bd_matrices[["predator_matrix"]]))
 
-#Bray Curtis Dissimilarity calculation
+#Bray Curtis Dissimilarity calculation----
+#by site
 BD_BySite <- bd %>%
   group_by(site) %>%
   summarise(across(MidgeFlies:RightHandedSnails, sum)) 
-  BC_BySite <- vegdist(select(BD_BySite, !site)) 
+  BC_BySite <- vegan::vegdist(select(BD_BySite, !site)) 
   BC_BySite_tibble <- tibble(
     site_comp = c("down-midstream", "down-upstream", "mid-upstream"),
     bc_score = as.vector(BC_BySite)
   )
   
-  
+  #by material
   BD_ByMat <- bd %>%
     group_by(material) %>%
     summarise(across(MidgeFlies:RightHandedSnails, sum)) 
-  BC_ByMat <- vegdist(select(BD_ByMat, !material)) 
+  BC_ByMat <- vegan::vegdist(select(BD_ByMat, !material)) 
   BC_ByMat_tibble <- tibble(
     mat_comp = c("Biopolymer-cellulose", "BioPoly-cotton", "Cell-cotton", "Biopoly-Jute", "Cell-jute", "Cotton-Jute", "Biopolymer-Plastic", "Cell-Plastic", "Cotton-Plastic", "Jute-plastic"),
     bc_score = as.vector(BC_ByMat)

### 2023-09-13 Rscripts/stats.R

@@ -28,6 +28,7 @@ diversity_tests <- diversity_mod_objects %>%
         shr_shannon = car::Anova(.$shr_shannon))}
 
 #perform tukey tests for all indices
+list() -> diversity_tukeys
 for(t in 1:length(diversity_tests)){
   
     cat(paste(names(diversity_mod_objects[t]), "had significant (<0.05) results.",

### 2023-09-13 Rdocs/ManuscriptLPP.Rmd
finished all plots!

@@
     geom_errorbar(aes(ymin = index - 2*index_se, ymax = index + 2*index_se)) +
-geom_text(aes(y = (index + 2*index_se)+0.05, label =groups)) +
+geom_text(aes(y = (index + 3*index_se), label =groups)) +
 
@@
   coord_cartesian(ylim = c(0, 1)) +
-  labs(title = "Community Dissimilairty between bag types", x = "Material Comparison",
+  labs(title = "Community Dissimilarity between plastic and biodegradable bags", x = "Material Comparison",
        y = "Bray Curtis Dissimilarity", fill = "Comparisons")
 
+#NMDS plots----
+#extract scores from NMDS object
+TC_NMDS_Scores <- as_tibble(scores(taxaCount_NMDS, display = "sites"))
+TC_NMDS_Scores$site <- LPP_FullData$site
+TC_NMDS_Scores$date <- LPP_FullData$date
+TC_NMDS_Scores$material <- LPP_FullData$material
+
+#plotting
+ggplot(TC_NMDS_Scores, aes(x = NMDS1, y = NMDS2, fill = material)) +
+  geom_hline(yintercept = 0, linetype = 5) + geom_vline(xintercept = 0, linetype = 5) +
+  geom_point(shape = 21, size = 2.5) +
+  facet_wrap(~date)+
+  theme_rafa() +
+  labs(title = "Non-Metric Dimensional Scaling of Community Matrices")+
+  scale_fill_viridis_d()
 ```

### 2023-09-13 Rscripts/stats.R

@@ -7,7 +7,7 @@ diversity_mod_objects <- list(#taxa wide anova model objects
   shannon = aov(shannon ~ material + as.factor(date) , data = LPP_FullData),
   simpson = aov(simpson ~ material + as.factor(date) , data = LPP_FullData),
   invsimp = aov(simpson ~ material + as.factor(date) , data = LPP_FullData),
-  evenness = aov(even ~ material + as.factor(date) , data = LPP_FullData),
+  even = aov(even ~ material + as.factor(date) , data = LPP_FullData),
   richness = aov(richness ~ material + as.factor(date), data = LPP_FullData),
   #functional group anova model objects
   coll_shannon = aov(collector_shannon ~ material + as.factor(date)   , data = LPP_FullData),
@@ -20,7 +20,7 @@ diversity_tests <- diversity_mod_objects %>%
   {list(shannon = car::Anova(.$shannon),
         simpson = car::Anova(.$simpson),
         invsimp = car::Anova(.$invsimp),
-        eveneness = car::Anova(.$evenness),
+        even = car::Anova(.$even),
         richness = car::Anova(.$richness),
         coll_shannon = car::Anova(.$coll_shannon),
         scr_shannon = car::Anova(.$scr_shannon),
@@ -73,5 +73,5 @@ set.seed(420)
 taxaCount_NMDS <- LPP_FullData %>%
   ungroup() %>%
   select(MidgeFlies:RightHandedSnails) %>%
-vegan::metaMDS(.,distance = "bray", k = 4, plot = T)
+vegan::metaMDS(.,distance = "bray", k = 3, plot = F)
 

### 2023-09-13 Rdocs/Plots.Rmd
finalized plot Rmd file, knitted to pdf output standardized plots

--- 2023-09-13 Rdocs/ManuscriptLPP.Rmd
@@
 ---
-title: Biodegradable alternatives to plastic mesh bags conventionally used in freshwater
-  macroinvertebrate analysis within the Bronx River, NYC
-author: "R.E. Segui-Baez, G.A. Rueda Moreno, R. Ward, K. Schneider Paolantonio"
-date: "`r Sys.Date()`"
-output: word_document
-bibliography: references.bib
+output:
+  pdf_document: default
+  html_document: default
 ---
 
-```{r pre-requisite scripts, echo=FALSE}
+```{r pre-requisite scripts, echo=FALSE, results='hide',message=FALSE,warning=FALSE}
+#set knitr options----
+knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
 #source data processor scripts----
@@
 
-```{r setup, include=FALSE}
-#knitr options----
-knitr::opts_chunk$set(echo = TRUE)
+```{r setup,echo=FALSE, results='hide',message=FALSE,warning=FALSE}
 #dependencies----
@@
     theme(axis.text.x = element_text(angle = -15, vjust = 0.5))+
-    labs(x = "material", title = "Diversity Index by Bag material", y = paste(index),
+    labs(x = "material", title = "Index by Bag material", y = paste(index),
          fill = "material")
@@
 
-### Tables
+# Leaf Pack Project Plots
 
-```{r wq table}
-readxl::read_xlsx("Rfigs/summaryFigures.xlsx", sheet = 2) %>%
-knitr::kable(caption = "Table 2: Water quality parameters averaged over the course of the sampling period")
-```
-
-### Plots
-
-```{r wq plots, eval=FALSE}
+```{r wq plots, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
 #water quality dot plots----
@@
   coord_cartesian(ylim = c(0, 1)) +
-  labs(title = "Community Dissimilarity between plastic and biodegradable bags", x = "Material Comparison",
+  labs(title = "Community Dissimilarity between \n plastic and biodegradable bags", x = "Material Comparison",
        y = "Bray Curtis Dissimilarity", fill = "Comparisons")
@@
   theme_rafa() +
-  labs(title = "Non-Metric Dimensional Scaling of Community Matrices")+
+  labs(title = "Bray Curtis NMDS of Community Matrices")+
   scale_fill_viridis_d()

### 2024-02-16 Rdocs/Plots.Rmd
saved wq plots as pngs edited wq processing script updated stats script with approrpiate normality and non-parametric tests added NMDS plot

@@
   #param(string): a string specifying the water quality parameter
-
 wq_ts_BySite <- function(param){
@@
   param_se <- sym(paste(param,"se", sep = "_"))
-  ggplot(wq_list$raw_water_quality, aes(x = date, y = !!param, color = site)) +
+  ggplot(wq_list$summ_water_quality, aes(x = date, y = !!param)) +
     #geom layers
-    geom_point() +
-     geom_smooth(method = 'lm')+
+    geom_jitter(height = 0, size = 2) +
+     geom_smooth(method = 'lm', se = T)+
     #labels
@@
     theme_rafa(base_size = 15) +
-    scale_color_viridis_d(begin = 0.1, end = 0.9)
+    scale_color_viridis_d()+
+  scale_fill_viridis_d()
+}
+
+wq_boxplot_BySite <- function(param){
+  #create symbol of parameter name and se for tidy eval in ggplot2 call
+  param<- sym(param)
+  param_se <- sym(paste(param,"se", sep = "_"))
+  ggplot(LPP_FullData, aes(x = site, y = !!param, fill = site)) +
+    #geom layers
+    geom_boxplot()+
+    geom_jitter(width = 0.5, height = 0.0, shape = 21, size = 2) +
+    #labels
+    labs(title = paste(param, "across sites"), x = "Site")+
+    #theme adjustments
+    theme_bw() +
+    theme_rafa(base_size = 15) +
+    scale_fill_viridis_d(option = "E",begin = 0.1, end = 0.9)
 }
@@
                             x = "Date", y = "Flow (m/s)")
+
+wq_boxplot_BySite("P") + labs(y = "Phosphorous (mg/L)", title = "Phosphorous Variation Across Sites")
+wq_boxplot_BySite("pH")
+wq_boxplot_BySite("NH3")+ labs(y = "Ammonia (mg/L)", title = "Ammonia Variation Across Sites")
+wq_boxplot_BySite("Nitrogen")+ labs(y = "Nitrogen (mg/L)", title = "Nitrogen Variation Across Sites")
+wq_boxplot_BySite("DO") + labs(y = " % DO")
+wq_boxplot_BySite("Flow") + labs(y = "Flow (m/s)", title = "Flow Rate Variation Across Sites")
+wq_boxplot_BySite("Cond") + labs(y = "Conductivity (mS)", title = "Conductivity Variation Across Sites")
+wq_boxplot_BySite("Temp") + labs(y = "Temp (°C)", title = "Temperature Variation Across Sites")
 
@@
 ```
+
+```{r}
+ggplot(wq_list$summ_water_quality) +
+  #geom_smooth(aes(x = date,color = site, y = NH3), method = "lm", se = T) +
+  geom_point(aes(x = date, fill =site, y = NH3), size = 2.5, shape = 21)+
+
+  scale_fill_viridis_d(option = "D") +
+  scale_color_viridis_d(option = "D")
+
+ggplot(wq_list$summ_water_quality, aes(x = NH3, y = site, fill = site)) +
+  geom_boxplot() +
+  geom_jitter(height = 0.2, width = 0.1, shape = 21, size = 1.5) +
+  theme_rafa(base_size = 12)+
+  facet_wrap(~date) +
+  scale_fill_viridis_d() +
+  labs(title = "")
+```
+
+```{r}
+#plot of first two NMDS dimensions, with ellipses for bag types
+ordiplot(taxaCount_NMDS)
+ordiellipse(taxaCount_NMDS,LPP_FullData$material, label = FALSE,
+            col=viridis(n = 5),
+            draw = "polygon", alpha=120) # adding ellipses to the plot, grouping by bag type
+title("Similaity of Communities Across Bag Types")
+legend("topright", title="Bag type",
+       c("Biopolymer", "Cellulose" , "Cotton",  "Jute", "Plastic" ), fill=viridis(n = 5), horiz=FALSE, cex=.9) # adding a legend
+```

### 2024-02-16 Rscripts/biodiversity.R

@@ -49,7 +49,7 @@ biodiversity <- bd %>%
 BD_BySite <- bd %>%
   group_by(site) %>%
   summarise(across(MidgeFlies:RightHandedSnails, sum)) 
-  BC_BySite <- vegan::vegdist(select(BD_BySite, !site)) 
+BC_BySite <- vegan::vegdist(select(BD_BySite, !site)) 
   BC_BySite_tibble <- tibble(
     site_comp = c("down-midstream", "down-upstream", "mid-upstream"),
     bc_score = as.vector(BC_BySite)

### 2024-02-16 Rscripts/stats.R

@@ -41,31 +41,46 @@ for(t in 1:length(diversity_tests)){
 }
 
 #WATER QUALITY MODELS----
+#normality tests
+wq_norms <- list(P = shapiro.test(wq_list$summ_water_quality$P),
+                 Nitrogen = shapiro.test(wq_list$summ_water_quality$Nitrogen),
+                 pH = shapiro.test(wq_list$summ_water_quality$pH),
+                 NH3 = shapiro.test(wq_list$summ_water_quality$NH3),
+                 DO = shapiro.test(wq_list$summ_water_quality$DO),
+                 Temp = shapiro.test(wq_list$summ_water_quality$Temp),
+                 Cond = shapiro.test(wq_list$summ_water_quality$Cond),
+                 FLow = shapiro.test(wq_list$summ_water_quality$Flow))
+
+
+#ANOVA for parameters which pass shapiro test
 #FORMULA: PARAMETER ~ DATE + SITE
 #wq model objects
-wq_mod_objects <- list(
-  P = aov(P~as.factor(date)+site, data = wq_list$raw_water_quality),
-  pH = aov(pH~as.factor(date)+site, data = wq_list$raw_water_quality),
-  NH3 = aov(NH3~as.factor(date)+site, data = wq_list$raw_water_quality),
-  DO = aov(DO~as.factor(date)+site, data = wq_list$raw_water_quality),
-  Temp = aov(Temp~as.factor(date)+site, data = wq_list$raw_water_quality),
-  Cond = aov(Cond~as.factor(date)+site, data = wq_list$raw_water_quality),
-  Flow = aov(Flow~as.factor(date)+site, data = wq_list$raw_water_quality)
+wq_normMod_objects <- list(
+  pH = aov(pH~site+date, data = wq_list$summ_water_quality),
+  Temp = aov(Temp~site+date, data =  wq_list$summ_water_quality),
+  Flow = aov(Flow~site+date, data =  wq_list$summ_water_quality)
 )
 
 #T2 anova
-wq_tests <- wq_mod_objects %>%
+wq_ANOVAt2 <- wq_normMod_objects %>%
   {list(
-  P = car::Anova(.$P),
   pH = car::Anova(.$pH),
-  NH3 = car::Anova(.$NH3),
-  DO = car::Anova(.$DO),
   Temp = car::Anova(.$Temp),
-  Cond = car::Anova(.$Cond),
   Flow = car::Anova(.$Flow))
   }
 
[…]

### 2024-02-16 Rscripts/water_quality.R

@@ -36,7 +36,8 @@ water_quality <- wq %>%
   pivot_longer(cols = ends_with("_Cond"), values_to = "Cond",names_to = "rn_Cond") %>%
   pivot_longer(cols = ends_with("_Nitrogen"), values_to = "Nitrogen",names_to = "rn_Nitrogen") %>%
   pivot_longer(cols = ends_with("_Flow"), values_to = "Flow",names_to = "rn_Flow") %>%
-  select(!starts_with("rn_"))
+  select(!starts_with("rn_")) %>% 
+  filter(P <4) #filter out outlier observations
 
 #summary statistics table grouped by site
 wq_site_summary <- water_quality %>%
@@ -49,6 +50,7 @@ wq_full_summary <- water_quality %>%
   group_by(site, date) %>%
   mutate(across(everything(), .fns = ~sd(., na.rm = T)/n(), .names = "{.col}_se")) %>%
   summarise(across(everything(), .fns = ~mean(., na.rm = TRUE)))
+
 
 #list of water quality objects
 wq_list <- list(raw_water_quality = water_quality,

### 2024-02-22 Rdocs/Plots.Rmd
Fixed the clusterfuck that was waterquality data, spaghetti code abounds
now The water quality plots are actually sensible, and thankfully stats remain unchanged.

sqrt transformed Nitrogen and Phosphorous, and now they are normal so I added them to the normal ANOVA group

Also, switched from doing independence test with coin and now im doing kruskal  wallis tests twice, once for site and once for date (not statistically responsible i know, but what choice do we have????)

I am dreading actually writing the manuscript

@@
   #create symbol of parameter name and se for tidy eval in ggplot2 call
-  param<- sym(param)
   param_se <- sym(paste(param,"se", sep = "_"))
-  ggplot(wq_list$summ_water_quality, aes(x = date, y = !!param)) +
+  param<- sym(paste(param, "mean", sep = "_"))
+
+  ggplot(water_quality, aes(x = date, y = !!param, color = site)) +
     #geom layers
-    geom_jitter(height = 0, size = 2) +
-     geom_smooth(method = 'lm', se = T)+
+    geom_errorbar(aes(ymin = !!param - 2*!!param_se, ymax = !!param +2*!!param_se), color = "black") +
+    geom_point( size = 2) +
+    facet_wrap(~site) +
     #labels
@@
   #create symbol of parameter name and se for tidy eval in ggplot2 call
-  param<- sym(param)
+
   param_se <- sym(paste(param,"se", sep = "_"))
+  param<- sym(paste(param, "mean", sep = "_"))
   ggplot(LPP_FullData, aes(x = site, y = !!param, fill = site)) +
@@
 ```{r}
-ggplot(wq_list$summ_water_quality) +
-  #geom_smooth(aes(x = date,color = site, y = NH3), method = "lm", se = T) +
-  geom_point(aes(x = date, fill =site, y = NH3), size = 2.5, shape = 21)+
-
-  scale_fill_viridis_d(option = "D") +
-  scale_color_viridis_d(option = "D")
-
-ggplot(wq_list$summ_water_quality, aes(x = NH3, y = site, fill = site)) +
-  geom_boxplot() +
-  geom_jitter(height = 0.2, width = 0.1, shape = 21, size = 1.5) +
-  theme_rafa(base_size = 12)+
-  facet_wrap(~date) +
-  scale_fill_viridis_d() +
-  labs(title = "")
+ggplot(water_quality, aes( x = sqrt(P_mean))) +geom_histogram(bins = 10)
+ggplot(water_quality, aes( x = Nitrogen_mean)) +geom_histogram(bins = 15)
 ```
@@
 ```
+
+```{r}
+#prop plots of functional groups
+FunGroups_props <- FunGroups %>% 
+  select(!FnG_Richness & !fnG_Shannon) %>% 
+  pivot_longer(cols = c("collector", "shredder", "scraper", "predator"), names_to = "fn_group", values_to = "counts") %>% 
+  group_by(material, date, site) %>% 
+  mutate(prop_counts = counts/sum(counts)) %>% 
+  group_by(material, fn_group) %>% 
+  mutate(avg_prop = mean(prop_counts),
+         se_prop = sd(prop_counts)/sqrt(n())) %>% 
+  summarise(avg_prop = unique(avg_prop),
+            se_prop = unique(se_prop))
+
+ggplot(FunGroups_props, aes(x = material, y = avg_prop, fill = fn_group)) +
+  geom_col(position = "dodge") +
+  geom_errorbar(aes(ymin = avg_prop -se_prop, ymax = avg_prop +se_prop), position = "dodge") +
+  scale_fill_viridis_d(option = "D", begin = 0.2) +
+  labs(title = "Distribution of Functional Groups across Bag types",
+       y = "Avg. Proportion", caption = "error bars show Standard Error margins")+
+  theme_rafa(base_size = 12)
+```

### 2024-02-22 Rscripts/biodiversity.R

@@ -37,7 +37,19 @@ bd_matrices <- list(
   predator_matrix = as.matrix(bd[predator])
 )
 
-#add functional group diversity as column
+#generate a cummul matrix of FunGroups
+FunGroups <- bd %>% 
+  group_by(site, material, date) %>% 
+  summarise(collector = sum(across(starts_with(collector))),
+            shredder = sum(across(starts_with(shredder))),
+            scraper = sum(across(starts_with(scraper))),
+            predator = sum(across(starts_with(predator)))) %>% 
+  ungroup() %>% 
+  mutate(FnG_Richness = hillR::hill_taxa(select(.,"collector":"predator"), q=0),
+         fnG_Shannon = hillR::hill_taxa(select(.,"collector":"predator"), q =1))
+  
+
+#add functional group diversity as column (within groups)
 biodiversity <- bd %>%
   mutate(collector_shannon = vegan::diversity(bd_matrices[["collector_matrix"]]),
          shredder_shannon = vegan::diversity(bd_matrices[["shredder_matrix"]]),

### 2024-02-22 Rscripts/stats.R

@@ -1,3 +1,6 @@
+#dependencies----
+require(coin)
+
 #join diversity data and water quality data together
 LPP_FullData <- right_join(x = wq_list$summ_water_quality, y = bd_list$biodiversity, by = c("date", "site"), multiple = "all")
 
@@ -42,43 +45,51 @@ for(t in 1:length(diversity_tests)){
 
 #WATER QUALITY MODELS----
 #normality tests
-wq_norms <- list(P = shapiro.test(wq_list$summ_water_quality$P),
-                 Nitrogen = shapiro.test(wq_list$summ_water_quality$Nitrogen),
-                 pH = shapiro.test(wq_list$summ_water_quality$pH),
-                 NH3 = shapiro.test(wq_list$summ_water_quality$NH3),
-                 DO = shapiro.test(wq_list$summ_water_quality$DO),
-                 Temp = shapiro.test(wq_list$summ_water_quality$Temp),
-                 Cond = shapiro.test(wq_list$summ_water_quality$Cond),
-                 FLow = shapiro.test(wq_list$summ_water_quality$Flow))
+wq_norms <- list(P = shapiro.test(sqrt(water_quality$P_mean)),
+                 Nitrogen = shapiro.test(sqrt(water_quality$Nitrogen_mean)),
+                 pH = shapiro.test(water_quality$pH_mean),
+                 NH3 = shapiro.test(water_quality$NH3_mean),
+                 DO = shapiro.test(water_quality$DO_mean),
+                 Temp = shapiro.test(water_quality$Temp_mean),
+                 Cond = shapiro.test(1/water_quality$Cond_mean),
+                 FLow = shapiro.test(water_quality$Flow_mean))
 
 
 #ANOVA for parameters which pass shapiro test
 #FORMULA: PARAMETER ~ DATE + SITE
 #wq model objects
 wq_normMod_objects <- list(
-  pH = aov(pH~site+date, data = wq_list$summ_water_quality),
-  Temp = aov(Temp~site+date, data =  wq_list$summ_water_quality),
-  Flow = aov(Flow~site+date, data =  wq_list$summ_water_quality)
+  sqrt_P = aov(sqrt(P_mean)~site+date, data = water_quality),
+  sqrt_Nitrogen = aov(sqrt(Nitrogen_mean)~site+date, data = water_quality),
[…]

### 2024-02-22 Rscripts/water_quality.R

@@ -5,6 +5,7 @@ wq <- readxl::read_excel("Rdata/LPP.xlsx",sheet = 2, na = "NA") %>%
   pivot_wider( names_from ="param", values_from = c("d_rep1", "d_rep2", "d_rep3",
                                                     "m_rep1", "m_rep2", "m_rep3",
                                                     "u_rep1", "u_rep2", "u_rep3")) #pivot wider  by the rep columns
+
 #separate into 3 tables corresp to site, and remove site prefix for easy merging
 #DOWNSTREAM
 d_wq <- wq %>%
@@ -25,37 +26,52 @@ u_wq <- wq %>%
 wq <- bind_rows(d_wq,m_wq,u_wq) %>%
   select(date, site, everything()) %>%
   arrange(date)
+  
 
 #pivot into longer format to condense rep columns into 1
 water_quality <- wq %>%
-  pivot_longer(cols = ends_with("_P"), values_to = "P",names_to = "rn_P") %>%
+  group_by(date, site) %>% 
+  pivot_longer(cols = ends_with("_P"), values_to = "P",names_to = "rn_P") %>% 
+  mutate(P_mean = mean(P, na.rm = T),
+         P_se = sd(P, na.rm = T)/sqrt(n())) %>% 
+  distinct(P_mean,.keep_all = T) %>% 
+  select(!rn_P) %>% 
   pivot_longer(cols = ends_with("_pH"), values_to = "pH",names_to = "rn_pH") %>%
+    mutate(pH_mean = mean(pH, na.rm = T),
+           pH_se = sd(pH, na.rm = T)/sqrt(n())) %>% 
+    distinct(pH_mean,.keep_all = T) %>% 
+  select(!rn_pH) %>% 
   pivot_longer(cols = ends_with("_NH3"), values_to = "NH3",names_to = "rn_NH3") %>%
+  mutate(NH3_mean = mean(NH3, na.rm = T),
+         NH3_se = sd(NH3, na.rm = T)/sqrt(n())) %>% 
+  distinct(NH3_mean,.keep_all = T) %>% 
+  select(!rn_NH3) %>% 
   pivot_longer(cols = ends_with("_DO"), values_to = "DO",names_to = "rn_DO") %>%
+  mutate(DO_mean = mean(DO, na.rm = T),
+         DO_se = sd(DO, na.rm = T)/sqrt(n())) %>% 
+  distinct(DO_mean,.keep_all = T) %>% 
[…]
